### 2023-03-12 outputs/paper/paper.qmd
upload paper and references as an example to later be updated with this same format

---
title: "Work Stress"
author: 
  - Sagith Kalaichelvam
  - Youngho Kim
thanks: "Code, data are available at: https://github.com/YoungKim164/workstress and https://doi.org/10.1257/jep.36.1.151 and replication package at: https://doi.org/10.48152/ssrp-cdf6-2r92"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "This paper explores several potential explanatory factors for the steep decline in birth rates in the United States between 2007 and 2020. The decline in U.S. birthrates offers some significant insights into U.S. society. Changes in birth patterns among recent young adult women moving through the childbearing age group can be directly related to the decline in birth rate from a purely statistical standpoint. We provide evidence that the steep drop in birth rates could be attributed to recent young adults' shifting interests. These shifting priorities represent changes in young adults' preferences regarding having children and aspirations for life. Resonating with the original paper, we conclude with a brief discussion of the implications of the declining birth rate in the United States and its impact on society as an entirety."
    toc: true
    number-sections: true
    colorlinks: true
    lot: true
    lof: true
bibliography: references.bib
---

```{r setup, include = FALSE}
#Load libraries
library(gridExtra)
library(cdlTools)
library(RGraphics)
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(ggplot2)
library(haven)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
```

# Introduction

To measure birth rates, the National Center for Health Statistics (NCHS) aggregates data on birth rates for women by age, race, education level, and other factors; the aggregated data for a particular year is the total number of births per 1000 women. Each year NCHS measures birth rates nationwide and analyzes the data to make insights into societal phenomena. For instance, in 1997, NCHS published a paper on how a mother's education level influences the birth rate. In the paper, NCHS asserted that there is a negative correlation between a mother's education level and birth rate. Their study found a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment @citeNCHS.

In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S. population data. With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR]. We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages. Figures and tables were created with knitr [@citeknit], ggplot2 [@citeggplot], kableExtra [@citekable], and dplyr [@citedplyr].

@fig-1 depicts U.S. birth rate trends over the last 40 years. Here, we can see that the trend is fluctuating relatively stable. However, during the post-2007 recession, it takes a steep decline with no sign of returning. We start our journey to explore potential explanations for the decline.

In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic. As shown in @fig-2, by dissecting the demographics into age groups and ethnicity groups, we notice a significant decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.

@sec-second-point, briefly discusses the empirical relationship between birth rates and economics and policy factors. We then argue that other economic, policy or social factors that have changed since 2007 do not contribute much to the decline.

@sec-third-point, presents a potential attributing factor for the declining U.S. birth rate after the 2007 recession. We show strong evidence that suggests the most recent group of young adults is a significant driver of the post-2007 decline in the birth rate. This significance suggests that the societal shifts among young women moving through childbearing age are an ongoing trend.

We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level. Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.

# Data {#sec-data}

## Data Source and Collection

The datasets for this paper were obtained from the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper], a paper from the Journal of Economic Perspectives [@citewebsite]. The data used by the paper was compiled from various sources such as the NCHS Natality Database [@citenchs], CDC SEER [@citecdcseer], the Current Population Survey Annual Social and Economic Supplement (CPS ASEC) [@citeipums], the New York Federal Reserve [@citefed], and the Integrated Public Use Microdata Series (IPUMS) [@citeipums]. The paper and datasets were published in the year 2022. These various sources provide a lot of useful information, shown throughout all of the figures we created in this paper. For example, the sources collected lots of raw data such as the population of people giving birth, race, lifestyle factors, age of parent when giving birth, etc. So, all of this information was collected from these various sources and written into the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper]. Which was then analyzed in our paper with our version of the analysis.

## Data Cleaning

Most of the data used in this paper were cleaned to get the most accurate information and remove any unnecessary information to be outputted in this paper. Data were cleaned to remove data of some years, age groups, race categories, etc. So that the revised figures can provide the most accurate information according to our analysis and not have any information that is unrelevant to our analysis. From our dataset, we dropped any values that are not related to the specific category, for example, values like NA. We have tried to filter our data much as we can for all of the graphs to make sure that it can make sense.

## Data Modification

We have included mainly all datasets and variables that are very important for our research question. We included variables by race, birth rates, age group, and other important factors. Lots of other variables were included, such as the total number of births and birth rates. Both of these allow us to analyze the decline of births over time. We did lots of modifications of the variables for this paper, which were mainly done within the age category. Generally, the original paper has an age category as one thing from age 15-44. For our figures, we made several figures and chose a specific age range. We have tried our best to divide the range into categories like 15-19, 20-24, 25-34, and 35-44 throughout most of the figures. We have also modified for race by having a figure that shows the data by different races, such as Black, White, and Hispanic people. Several figures had an important modifications to be relevant to our analysis throughout the paper.

## Data Visualization

Most of the figures show lots of visualization of the data. We can easily expect that the birth rate is declining over time within the United States due to many other factors impacting people's decision to have more kids. When it comes to race, there are a lot of data showing some useful information to visualize. Overall, black people tend to have lower birth rates than Hispanic and white people. Hispanics have the highest birth rates, while people are the second highest. But all races are showing a down trendline that the birth rates are going down over time which can be coming from other factors. When it comes to factors like childcare expenses, it shows that people with lower expense tends to have low birth rates, while more expense tends to have more birth rates. When the rental housing costs are high, the number of birth rates tends to go down, and the same for the situation in student loan debt. When it comes to the figures of different races, we decided to focus on age groups such as 15-19 and 20-24 because we believe that this age category has more useful information since we know that in the past, people tended to have kids much earlier than before due to having marriage at an early age. The birth rates by different races in the age group 15-19 have drastically decreased over time when compared to the age group 20-24. This is because fewer people are getting married before age 20 and having kids before age 20.

# Results {#sec-results}

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-1
#| fig-cap: Trends in U.S. Birth rates of 1000 women between age 15 to 44 from 1980 to 2020

fig1_data <- read.csv(here::here("inputs/data/fig_1.csv"))
fig1_data %>%
  ggplot(aes(x = year,
             y = brate_all)) +
  geom_line() +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  labs(
    title = "",
    x = "Years",
    y = "Births per 1,000 women age 15-44",
    caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019 and 2020."
  )  +  
  geom_vline(xintercept = 2007, linetype = "dashed") +
  geom_text(aes(x = 2007, y = 70), nudge_x = 2, nudge_y = 1, label = "2007") +
  theme_light()
```

@fig-1 shows that over time, the number of births in the United States is dropping steadily, while it slightly went up on peak around 1990, 2001, and 2007. Most likely, the hikes going up during these years might have been into other factors that made more people among the selected population have kids. After 2007, the number of births drastically dropped a lot. Nowadays, among 1,000 women from age 15 to 44, only about 56 births will occur. While in the past, around the 1980s and 1990s, the number of births was around the 65 to 70 range. The line in the graph shows that in the near future, there will be a much lower amount of people having kids, and this can be due to other factors from their lifestyle, which is making them decide not to have more kids.

While the figure shows the birth rate trend is going down, there are certain spikes in the peak of birth rates around 1990, 2001, and 2007. There were some incidents that must have happened in the United States in those years, which may have caused people to stay home more often. As we know that people that are likely to be unemployed tend to have more kids as they will have more freedom during their time spent at home. This allows the birth rates to be higher than usual. During these years, Americans faced lower employment opportunities within society [@citecatholic]. They do not have access to useful resources that improve their lifestyle in these years, so they are at home most of the time, which encourages couples to have an interest in having more kids, which ends up increasing the birth rates for these years.

\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-2
#| fig-cap: Trend in Birthrate by Race and Ethnicity for Ages 15-19 and 20-24 from 2000 to 2020)
birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

# getting total number of births for white people from 15 to 19 years old
birth1$numbirthwhite15 = birth1$numbirth_lesshs_whitenh_1519 + birth1$numbirth_educ_miss_whitenh_1519 + birth1$numbirth_gtc_whitenh_1519 + birth1$numbirth_somecoll_whitenh_1519 + birth1$numbirth_hsgrad_whitenh_1519 + birth1$numbirth_ba_whitenh_1519

# getting total number of births for black people from 15 to 19 years old  
birth1$numbirthblack15 = birth1$numbirth_lesshs_blacknh_1519 + birth1$numbirth_educ_miss_blacknh_1519 + birth1$numbirth_gtc_blacknh_1519 + birth1$numbirth_somecoll_blacknh_1519 + birth1$numbirth_hsgrad_blacknh_1519 + birth1$numbirth_ba_blacknh_1519

# getting total number of births for hispanics from 15 to 19 years old  
birth1$numbirthhisp15 = birth1$numbirth_lesshs_hisp_1519 + birth1$numbirth_educ_miss_hisp_1519 + birth1$numbirth_gtc_hisp_1519 + birth1$numbirth_somecoll_hisp_1519 + birth1$numbirth_hsgrad_hisp_1519 + birth1$numbirth_ba_hisp_1519

# getting total number of births for white people from 20 to 24 years old
birth1$numbirthwhite20 = birth1$numbirth_lesshs_whitenh_2024 + birth1$numbirth_educ_miss_whitenh_2024 + birth1$numbirth_gtc_whitenh_2024 + birth1$numbirth_somecoll_whitenh_2024 + birth1$numbirth_hsgrad_whitenh_2024 + birth1$numbirth_ba_whitenh_2024

# getting total number of births for black people from 20 to 24 years old
birth1$numbirthblack20 = birth1$numbirth_lesshs_blacknh_2024 + birth1$numbirth_educ_miss_blacknh_2024 + birth1$numbirth_gtc_blacknh_2024 + birth1$numbirth_somecoll_blacknh_2024 + birth1$numbirth_hsgrad_blacknh_2024 + birth1$numbirth_ba_blacknh_2024

# getting total number of births for hispanics from 20 to 24 years old  
birth1$numbirthhisp20 = birth1$numbirth_lesshs_hisp_2024 + birth1$numbirth_educ_miss_hisp_2024 + birth1$numbirth_gtc_hisp_2024 + birth1$numbirth_somecoll_hisp_2024 + birth1$numbirth_hsgrad_hisp_2024 + birth1$numbirth_ba_hisp_2024

# gets the total number of births by year, compiled nationally
birthhist = birth1 %>%
  filter(year >= 2000) %>%
  group_by(year) %>% 
  select(year,pop1544, numbirthwhite15,numbirthwhite20,numbirthhisp15,numbirthhisp20,numbirthblack15,numbirthblack20) %>%
  summarize(sumpop = sum(pop1544),white15 = sum(numbirthwhite15), white20 = sum(numbirthwhite20), black15 = sum(numbirthblack15), black20 = sum(numbirthblack20), hisp15 = sum(numbirthhisp15), hisp20 = sum(numbirthhisp20))

# gets the birth rate for each race by adding the number of births from the 15 to 19for each race, then multiplying by 10,000 and dividing by the female population
fig21<-birthhist %>%
  mutate(whiteall = white15*10000/sumpop, blackall= black15*10000/sumpop, hispall = hisp15*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=4))+
  labs(title = "Age Group 15 to 19",
       x = "Year",
       y = "Birth Rate",
       color = "Race") +
  scale_color_brewer(palette = "Set1", labels = c("Black", "Hispanic", "White")) +
  theme(legend.position = "bottom")

fig22<-birthhist %>%
  mutate(whiteall = white20*10000/sumpop, blackall= black20*10000/sumpop, hispall = hisp20*10000/sumpop) %>% 
  select(year, whiteall, blackall, hispall) %>%
  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
               names_to = "racegroup",
               values_to = "value"
  ) %>%
  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
  geom_line() +
  theme_minimal() +
  scale_x_continuous(breaks = seq(2000,2020, by=4))+
  labs(title="Age Group 20 to 24",
       x = "Year",
       y = "Birth Rate",
       color = "Race",
       caption = "Source: Birth Rates collected from
                  CDC Vital Statistics Births Reports
                  for 2015, 2019 and 2020.") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "none")

grid.arrange(fig21,fig22,ncol=2,nrow=1)

```

@fig-2 shows the trend in birth rates over the period 2000-2019 in different races for two different age groups. In all race group, the birth rates starts to drop significantly after the Great Recession. In a age group of 15-19, all three race groups had similar drops on the birth rate. On the other hand in age group 20-24 the white race group showed significanat birth rate drop from 86 to 49 birth per 1000 population., where as other race groups had much less birthrate drop.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-4
#| fig-cap: Changes in Other Economic factors and Birth Rate over the periods of 2004-2008 and 2014-2019

#04~19 birthrate data
birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
pop1 = read_dta(here::here("inputs/data/pop/age_race_comp_seer.dta"))
birth1 <- birth1 %>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019))

birth1 <- birth1 %>% 
  subset(select=c(stname,year,numbirth1544))

pop1<-pop1 %>% 
  mutate(pop1544 = pop1519 + pop2034 + pop3544)%>% 
  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019)) %>%
  group_by(year,stname) %>%
  select(pop1544) %>% 
  summarise(across(everything(),sum))

fig4_data <- right_join(birth1, pop1, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  mutate(year = case_when(year=2004 & year<=2008 ~ 2004,
                           year>=2015 & year<=2019 ~ 2019)) %>%
  group_by(year,stname) %>%
  select(numbirth1544,pop1544) %>% 
  summarise(across(everything(),sum)) %>% 
  mutate(birthrate=numbirth1544/pop1544*1000) %>% 
  select(-c(numbirth1544,pop1544))

fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(birthrate = birthrate[year == 2019] - birthrate[year == 2004])

#04~19 childcare cost for fig4b
childcare_aesc=read_dta(here::here("inputs/data/long_term/childcare_asec_09_1519.dta"))
childcare_aesc <- childcare_aesc %>% 
  mutate(childcare_cost=child_asec)
childcare_aesc <- childcare_aesc %>% 
  select(year,stname,childcare_cost)
fig4_data <- right_join(fig4_data, childcare_aesc, by=c("stname","year"))
fig4_data <- fig4_data %>% 
  group_by(stname) %>% 
  mutate(childcare_cost = childcare_cost[year == 2019] - childcare_cost[year == 2004])

#04~19 rentcost data for fig4c
rentcost_0419=read_dta(here::here("inputs/data/long_term/longdiff-rentcost-0419.dta"))

rentcost_0419 <- rentcost_0419 %>% 
  mutate(stname = fips(rentcost_0419$statefip, to='Abbreviation')) %>%
  group_by(stname) %>% 
  mutate(avg_grossrent = avg_grossrent[year == 2019] - avg_grossrent[year == 2004]) %>% 
  select(year,stname,avg_grossrent)

fig4_data <- right_join(fig4_data, rentcost_0419, by=c("stname","year"))

#04~19 wage ratio data for fig4d
rhs_1519=read_dta(here::here("inputs/data/long_term/longdiff-RHS-1519.dta"))
rhs_0408=read_dta(here::here("inputs/data/long_term/longdiff-RHS-0408.dta"))
rhs<-rbind(rhs_1519, rhs_0408)
rhs <- rhs %>% 
  mutate(stname = fips(rhs$statefip, to='Abbreviation')) %>% 
  mutate(fem_male_ratio=p50wage_all/p50wage_male_all) %>% 
  select(year,stname,fem_male_ratio) %>% 
  group_by(stname) %>% 
  mutate(fem_male_ratio = fem_male_ratio[year == 2019] - fem_male_ratio[year == 2004])

fig4_data <- right_join(fig4_data, rhs, by=c("stname","year"))

#04~19 student loan data for fig4e
std_loan=read_dta(here::here("inputs/data/long_term/nyfed-student-debt-per-capita-0419.dta"))

std_loan<- std_loan %>% 
  group_by(stname) %>% 
  mutate(stud_loan_cap = stud_loan_cap[year == 2019] - stud_loan_cap[year == 2004])

fig4_data <- right_join(fig4_data, std_loan, by=c("stname","year"))

fig4_data <- fig4_data %>% 
  head(51) %>% 
  select(-c(year))

#drawing
fig4b<-fig4_data %>% 
  ggplot(aes(x=childcare_cost*1000, y=birthrate)) +
  geom_point(size=0)+
  labs(
    title = "Childcare Expenditures",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 5000), breaks=seq(0, 5000, by = 1000),labels = paste0("$",seq(0, 5000, by=1000))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4c <-fig4_data %>% 
  ggplot(aes(x=avg_grossrent, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Rental Housing Costs",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(0, 400), breaks=seq(0, 400, by = 75),labels = paste0("$",seq(0, 400, by = 75))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4d<-fig4_data %>% 
  ggplot(aes(x=fem_male_ratio, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Female-Male Wage Ratio",
    x = "Fem to Male Wage Ratio",
    y = "Birthrate")+
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(-0.08, 0.12), breaks=seq(-0.08, 0.12, by = 0.04)) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

fig4e<-fig4_data %>% 
  ggplot(aes(x=stud_loan_cap*1000, y=birthrate)) +
  geom_point(size=0)  +
  labs(
    title = "Student Loan Debt",
    x = "US dollar",
    y = "Birthrate") +
  geom_text(aes(label = stname),size=2) +
  scale_x_continuous(limits=c(1000, 4500), breaks=seq(1000, 4500, by = 750),labels = paste0("$",seq(1000, 4500, by=750))) +
  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
  geom_smooth(method="lm",se=FALSE, size = 0)  +
  theme_light()

grid.arrange(fig4b,fig4c,fig4d,fig4e,ncol=2,nrow=2)

```

@fig-4 shows an interesting relationship between changes in birthrate and other economic factors over five years period. Each marker represents a state labelled with its two-letter abbreviation.

Annual expenditures on childcare for families with children under 12 who report positive childcare spending rose \$2,170 nationwide for the average of 2015--2019, based on data from the Current Population Survey (all dollar values throughout are measured in constant 2019 dollars). But in the cross-state correlations, places where childcare expenditures increased more, did not experience a noticeable drop in birth rates. In fact, the trend shows that more increases in childcare expenditure have no significant impact on birthrate changes.

Over this period, average monthly rents for a two- to three-bedroom apartment rose \$124 per month (from \$930 to \$1,060 in 2018 dollars, a 14 percent increase) nationwide. The increase was much larger in some states, like Colorado, Washington, and the District of Columbia. The data, however, do not indicate a negative association between rent changes and birth rate changes.

We used the female-to-male median earnings ratio among full-time, year-round workers to examine women's economic opportunities. On average, the median earning ratio rose to 0.04 in the period. 12 states where the wage ratio rose the most over this sample period did not show a greater birth rate decline.

If young adults are saddled with debt, they might not have sufficient disposable income to have a child or more children. We consider the total student debt per capita level in a state, which has increased from \$2,500 to \$5,400 (in 2018 dollars) on average. The relationship between state-level student debt and the birth rate is generally flat. It indicates that increases in student debt are related to aggregate reductions in birth rates.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-5
#| fig-cap: Trends in Birth Rates by Age Groups from 1990-2019

# Reading in the data.

birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))

birth1 <- birth1 %>%
  group_by(year) %>%
  select(-c(stname)) %>% 
  summarise(across(everything(),sum))

birth1 <- birth1 %>% 
  mutate(birthrate1519 = numbirth1519/pop1519*1000) %>% 
  mutate(birthrate2024 = numbirth2024/pop2024*1000) %>% 
  mutate(birthrate2529 = numbirth2529/(pop25+pop26+pop27+pop28+pop29)*1000) %>% 
  mutate(birthrate3034 = numbirth3034/(pop30+pop31+pop32+pop33+pop34)*1000) %>% 
  mutate(birthrate3539 = numbirth3539/(pop35+pop36+pop37+pop38+pop39)*1000) %>% 
  mutate(birthrate4044 = numbirth4044/(pop40+pop41+pop42+pop43+pop44)*1000)

birth1 <- birth1 %>% 
  select(year, birthrate1519, birthrate2024, birthrate2529, birthrate3034, birthrate3539, birthrate4044) %>%
  pivot_longer(-c(year),
               names_to = "agegroup",
               values_to = "birthrate"
  )

birth1<-birth1 %>% 
  mutate(agegroup = as.character(str_replace(agegroup, "birthrate", "")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "1519", "15~19")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2024", "20~24")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2529", "25~29")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3034", "30~34")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3539", "35~39")))
birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "4044", "40~44")))

birth1 %>% 
ggplot(mapping = aes(x = year, y = birthrate, color = agegroup)) + 
  geom_line()  +
  labs(
    title = "",
    x = "Year",
    y = "birthrate",
    caption = "Source: CDC Vital Statistics Births Reports.",
    color = "Age Group"
  ) +
  scale_x_continuous(limits=c(1991, 2020), breaks=seq(1991, 2020, by = 8)) +
  scale_y_continuous(limits = c(0,125), breaks = seq(0,125, by=20)) +
  theme_light()
```

@fig-5 shows lots of important information about when people tend to have kids at certain age over the timeframe from 1990 to 2020. Based on the graph, the data shows that more people have kids after the age of 30 during this present time. In the past, most people tend to have kids before the age of 30. Around 2000 to 2005, the lifestyle started to change, where people tended to either get married or have kids later, which is why the trend of birth rates in age 30 and up has been going higher since the 2000s. While births before the age of 30 tend to slow down and decrease over time from the 2000s. According to the graph, maybe in another 100 years, we can expect that most birth rates can happen around 40-44.

There are drastic changes shown in @fig-5, such as age 30-44 tends to increase around the 2000s, while the ages below 30 decreases for the number of birth rates. Around the decade of the 2000s or the beginning of this new 21st century, there was lots of modernization from people's lifestyle to other things on Earth, which is coming from several factors such as education and employment due to new innovations and inventions [@citedimock]. There are lots of new behaviour being introduced to a new generation of people. So, as they get new behaviour of attitudes, that is being involved in their decisions of marriages, and when to have kids. This is why lots of people are now getting married late and/or having kids much later when compared to the past.

Figure 5. Children Ever Born by Mother's Age and Birth Cohort (@fig-6).

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-6
#| fig-cap: Children Ever Born by Mother's Age and Birth Cohort

nch_cohot_data <- read_dta(here::here("inputs/data/nchs/nchs_cohort_analysis.dta"))
agecomp_data <- read_dta(here::here("inputs/data/pop/agecomp-seer.dta"))

pop_by_cohort<- agecomp_data %>%
  select(1:32) %>%
  group_by(year) %>%
  select(-c(stname)) %>%
  summarise(across(everything(),sum)) %>%
  pivot_longer(-c(year), names_to = "mage", values_to = "pop") %>%
  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
  mutate(birthyear = year - mage) %>%
  mutate(cohort2 = case_when(birthyear >= 1968 & birthyear <=1972 ~ 1,
                            birthyear >= 1973 & birthyear <=1977 ~ 2,
                            birthyear >= 1978 & birthyear <=1982 ~ 3,
                            birthyear >= 1983 & birthyear <=1987 ~ 4,
                            birthyear >= 1988 & birthyear <=1992 ~ 5,
                            birthyear >= 1993 & birthyear <=1997 ~ 6)) %>% 
  group_by(cohort2,mage) %>%
  summarise(pop = sum(pop))

fig5_data <- right_join(nch_cohot_data, pop_by_cohort, by=c("mage", "cohort2"))
fig5_data <- fig5_data %>%
  mutate(birthrate=numbirth/pop*1000,
         age_2024_year = case_when(cohort2 == 1 ~ 1992,
                                   cohort2 == 2 ~ 1997,
                                   cohort2 == 3 ~ 2002,
                                   cohort2 == 4 ~ 2007,
                                   cohort2 == 5 ~ 2012,
                                   cohort2 == 6 ~ 2017))

fig5_data <- fig5_data %>% 
  group_by(age_2024_year) %>% 
  mutate(cum_birthrate = cumsum(birthrate)/1000, 
         cohort = case_when(cohort2 == 1 ~ "1968-1972        1992",
                                  cohort2 == 2 ~ "1973-1977        1997",
                                  cohort2 == 3 ~ "1978-1982        2002",
                                  cohort2 == 4 ~ "1983-1987        2007",
                                  cohort2 == 5 ~ "1988-1992        2012",
                                  cohort2 == 6 ~ "1993-1997        2017"))

fig5_data <-  na.omit(fig5_data)
fig5_data %>%
  ggplot(aes(x = mage,
             y = cum_birthrate,
             color = cohort)) +
  geom_line()  +
  labs(
    title = "",
    x = "Ages",
    y = "Cumulative birthrate",
    color = "Birth cohorts   Age 20 in"
  ) +
  scale_x_continuous(limits=c(15, 44), breaks=seq(15, 44, by = 3)) +
  theme_minimal()+
  theme(legend.position = "right")

```

@fig-6 shows lots of useful information about the cumulative birth rate by mother's age in several birth cohorts from early 1968 to late 1997. We tend to use these years since, in recent times, we believe that there is enough data for these years since most of the births and living people are among these cohorts. According to @fig-6, the earlier the cohort is, the cumulative birth rate will be really high. While the later the cohort is, the cumulative birth rate will be much lower. The highest cumulative birth rate is coming from the birth cohort of 1968-1972 and 1973-1977. While the lowest comes from the 1993-1997 birth cohort.
\newpage

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-1
#| tbl-cap: The Eight Demographic Groups that Contributed the Most to the 2007-2019 Decline in the U.S. Brith Rate
numbirth_2044 <- read.csv(here::here("inputs/data/edu/numbirths_educ_2044.csv"))
numbirth_2044 %>% 
  kable(col.name=c("Age Group","Change in Birth Rate"),
        booktabs = TRUE,show_col_types=FALSE,linesep="",
        title = "The Eight Demographic Groups that Contributed to the Most to the 2007-2019 Decline in the U.S. Brith Rate"
        )

```

@tbl-1 shows the most contributor to the decline in birth rates across all age groups. The top contributor is Hispanic with less than high school education women between age 20-24 at -88.3, Hispanic with high school education women between age 15-19 at -57.5, and Black (Non-Hispanic) women between age 15-19 at -39.8, respectively.

\newpage

# Discussion

## Birth Rates by Demographic Groups {#sec-first-point}

In 1997, NCHS released a report titled "Mother's Educational Level Influences Birth Rate." According to the report, a women's education level is the best predictor of how many children she will have. In the report, NCHS describes a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment [@citeNCHS].\
We are particularly interested in ethnic groups of ages 15-19 and 20-24. @fig-2 shows a sharp decline in birth rates among all groups of women between the age of 15 to 19 during the post-2007 recession. @fig-2 also shows a sharp decline over the same period; however, only among white ethnic women between the ages 20 to 24, while other groups show a relatively steady decline. Figures that are omitted from this paper of the other age groups (25 to 34 and 35 to 44) show a similar trend both in pre and post-recession. However, as shown in @tbl-1, the most significant contributor to the decline in the birth rate is Hispanic, with less than High School education, women between ages 20-24 after 2007. This result contradicts the earlier claim made by NCHS. The finding suggests, in the post-2007 recession, the women's education level and birth rate now correlate positively.

## Economic and Policy Factors {#sec-second-point}

In @fig-4, we show other economic factors that might have negatively affected the birth rate in the post-2007 recession across the states. Although the graphs show their inverse relationship, no dramatic change is shown. For instance, as there were somewhere around \$1000 to \$2000 increases in childcare expenses over the past 13 years, we see that the decrease in the actual birth rate is only between 0 to -10. Change in per capita student loans also has increased between \$2500 to \$3250, but it only contributed to the decrease somewhere between 0 to -5 across the states. Based on these findings, it is hard to find an explanation of the dramatic decline in the birth rate since 2007 from economic factors.

## A Potential Explanation: Shifting Priorities {#sec-third-point}

When it comes to the age of the population of the United States and comparing it from the past to now. It shows that more people tend to have kids much later than before. While fewer people have kids earlier than in the past. In the past, more people tend to have kids much earlier. There are several reasons why more people had kids at an early age back then; people used to get married much earlier, and many people were unemployed or doing a job for self-employment, which was not even an educated or professional job. The modern lifestyle that has been adopted in the past few decades tends to make people focus on all other areas of life, which makes them marry much later, and have kids later. Nowadays, many marriages happen at the age of late 20s or early 30s, while they tend to have their first child around the early to mid-30s. @fig-5 depicts the birth rates in women between the ages 30-39 and 35-39 positive slope trend. Whether pre or post-2007 recession, this trend looks steady over the past 40 years. This finding aligns with the current common belief and the social norm where, perhaps, women's peak childbearing ages have shifted from the early twenties to early thirties. @fig-6 shows the average number of children born to women by age in five-year intervals from the 1968-1972 through 1993-1997 birth cohorts. These birth cohorts include women entering their peak childbearing years (20--24) in 1992, 1997, 2002, 2007, and 2017. This chart indicates that the three cohorts of young adult women born between 1968 and 1982 in 1992, 1997, and 2002 had similar childbearing age profiles. The 2007 birth cohort of women had fewer children in their 20s and 30s. The 2012 and 2017 birth cohorts (born between 1988 and 1997) have significantly fewer children than the preceding cohorts. This finding is consistent with a decline in births, primarily to more recent cohorts of women. From a statistical point of view, this divergence across cohorts explains the sizable decrease in annual birth rates that began in 2007.

## Weakness and Moving Forward

### Bias and Ethical Concerns

As we know that this data was collected and created by other human beings, just as how we are. There are some ethical concerns that could affect the accuracy of our data because humans can make mistakes, and they can make decisions based on their preferences, which can affect the whole dataset. So, this analysis is only valid based on our dataset, which can be modified if the original dataset is changed.

The original data shows different races, such as White, Black, and Hispanic. They have included Hispanic based on all colours, which erases the identity of Black and Indigenous Hispanic people. This tends to be a bit of racism because the information provided on the data did not specifically indicate whether or not the Hispanic category includes everyone that identifies as Hispanic or is it just people that look Hispanic.

### Data Collection Concerns

This dataset tends to collect huge data information on the population among people from the United States and considers lots of factors from race, birth rates by age, student debts, etc. Having too much information means that when the data is collected, there is a possibility of human error, calculation error, and chances of small modification of data due format of reporting an error. So, the way how the data was collected can have lots of problems because people can make mistakes easily.

### Moving Forward

This paper raised concerns about a dramatic decline in U.S. birth rates starting in 2007. Through the paper, we proposed a general explanation for the decline in births across recent cohorts of U.S. women that focuses on the shifting priorities of cohorts. As dictated by the recent socio-phenomena that encompasses preferences for having children, life aspirations, and the nature of parenting, a change in the relationship between education and childbearing, a rise in childlessness, and the establishment of a two-child norm for those having children have irreversibly changed [@citepaper]. This new social phenomenon that might be causing the dramatic decline in U.S. birth rates is important from an economic policy standpoint to acknowledge that an aging population and shrinking workforce pose challenges to economic growth and the sustainability of social insurance systems [@Yang2003]. Thus, the most appropriate way to address declining U.S. birth rates may be required: 1. developing A.I. and robots where a shrinking workforce poses a problem, 2. an appropriate policy to address issues with Social Security and Medicare on a secure basis for the long-term. Suppose the U.S. birth rate is not showing any reverse trend shortly. In that case, the U.S. economy and political system should enforce greater emphasis on addressing the issue before it becomes irreversible.

\newpage

# References

### 2023-03-13 outputs/paper/paper.qmd
Update paper to include the correct abstract

@@
     geometry: margin = 1in
-    abstract: "This paper explores several potential explanatory factors for the steep decline in birth rates in the United States between 2007 and 2020.
-The decline in U.S.
-birthrates offers some significant insights into U.S.
-society.
-Changes in birth patterns among recent young adult women moving through the childbearing age group can be directly related to the decline in birth rate from a purely statistical standpoint.
-We provide evidence that the steep drop in birth rates could be attributed to recent young adults' shifting interests.
-These shifting priorities represent changes in young adults' preferences regarding having children and aspirations for life.
-Resonating with the original paper, we conclude with a brief discussion of the implications of the declining birth rate in the United States and its impact on society as an entirety."
+    abstract: "This paper discusses data from some survey, which can gather data that is linked to how people face work stress.
+This paper uses the original data from the survey which determines the relationships between respondent work stress levels by satisfaction, status, sectors, and work time.
+This paper will provide a strong analysis about the work stress that people are facing."
     toc: true

### 2023-03-13 outputs/paper/paper.qmd
Update paper to remove unrelated figures

@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-1
-#| fig-cap: Trends in U.S. Birth rates of 1000 women between age 15 to 44 from 1980 to 2020
-
-fig1_data <- read.csv(here::here("inputs/data/fig_1.csv"))
-fig1_data %>%
-  ggplot(aes(x = year,
-             y = brate_all)) +
-  geom_line() +
-  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
-  labs(
-    title = "",
-    x = "Years",
-    y = "Births per 1,000 women age 15-44",
-    caption = "Source: Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019 and 2020."
-  )  +  
-  geom_vline(xintercept = 2007, linetype = "dashed") +
-  geom_text(aes(x = 2007, y = 70), nudge_x = 2, nudge_y = 1, label = "2007") +
-  theme_light()
-```
-
 @fig-1 shows that over time, the number of births in the United States is dropping steadily, while it slightly went up on peak around 1990, 2001, and 2007.
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-2
-#| fig-cap: Trend in Birthrate by Race and Ethnicity for Ages 15-19 and 20-24 from 2000 to 2020)
-birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
-
-birth1 <- birth1 %>%
-  group_by(year) %>%
-  select(-c(stname)) %>% 
-  summarise(across(everything(),sum))
-
-# getting total number of births for white people from 15 to 19 years old
-birth1$numbirthwhite15 = birth1$numbirth_lesshs_whitenh_1519 + birth1$numbirth_educ_miss_whitenh_1519 + birth1$numbirth_gtc_whitenh_1519 + birth1$numbirth_somecoll_whitenh_1519 + birth1$numbirth_hsgrad_whitenh_1519 + birth1$numbirth_ba_whitenh_1519
-
-# getting total number of births for black people from 15 to 19 years old  
-birth1$numbirthblack15 = birth1$numbirth_lesshs_blacknh_1519 + birth1$numbirth_educ_miss_blacknh_1519 + birth1$numbirth_gtc_blacknh_1519 + birth1$numbirth_somecoll_blacknh_1519 + birth1$numbirth_hsgrad_blacknh_1519 + birth1$numbirth_ba_blacknh_1519
-
-# getting total number of births for hispanics from 15 to 19 years old  
-birth1$numbirthhisp15 = birth1$numbirth_lesshs_hisp_1519 + birth1$numbirth_educ_miss_hisp_1519 + birth1$numbirth_gtc_hisp_1519 + birth1$numbirth_somecoll_hisp_1519 + birth1$numbirth_hsgrad_hisp_1519 + birth1$numbirth_ba_hisp_1519
-
-# getting total number of births for white people from 20 to 24 years old
-birth1$numbirthwhite20 = birth1$numbirth_lesshs_whitenh_2024 + birth1$numbirth_educ_miss_whitenh_2024 + birth1$numbirth_gtc_whitenh_2024 + birth1$numbirth_somecoll_whitenh_2024 + birth1$numbirth_hsgrad_whitenh_2024 + birth1$numbirth_ba_whitenh_2024
-
-# getting total number of births for black people from 20 to 24 years old
-birth1$numbirthblack20 = birth1$numbirth_lesshs_blacknh_2024 + birth1$numbirth_educ_miss_blacknh_2024 + birth1$numbirth_gtc_blacknh_2024 + birth1$numbirth_somecoll_blacknh_2024 + birth1$numbirth_hsgrad_blacknh_2024 + birth1$numbirth_ba_blacknh_2024
-
-# getting total number of births for hispanics from 20 to 24 years old  
-birth1$numbirthhisp20 = birth1$numbirth_lesshs_hisp_2024 + birth1$numbirth_educ_miss_hisp_2024 + birth1$numbirth_gtc_hisp_2024 + birth1$numbirth_somecoll_hisp_2024 + birth1$numbirth_hsgrad_hisp_2024 + birth1$numbirth_ba_hisp_2024
-
-# gets the total number of births by year, compiled nationally
-birthhist = birth1 %>%
-  filter(year >= 2000) %>%
-  group_by(year) %>% 
-  select(year,pop1544, numbirthwhite15,numbirthwhite20,numbirthhisp15,numbirthhisp20,numbirthblack15,numbirthblack20) %>%
-  summarize(sumpop = sum(pop1544),white15 = sum(numbirthwhite15), white20 = sum(numbirthwhite20), black15 = sum(numbirthblack15), black20 = sum(numbirthblack20), hisp15 = sum(numbirthhisp15), hisp20 = sum(numbirthhisp20))
-
-# gets the birth rate for each race by adding the number of births from the 15 to 19for each race, then multiplying by 10,000 and dividing by the female population
-fig21<-birthhist %>%
-  mutate(whiteall = white15*10000/sumpop, blackall= black15*10000/sumpop, hispall = hisp15*10000/sumpop) %>% 
-  select(year, whiteall, blackall, hispall) %>%
-  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
-               names_to = "racegroup",
-               values_to = "value"
-  ) %>%
-  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
-  geom_line() +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(2000,2020, by=4))+
-  labs(title = "Age Group 15 to 19",
-       x = "Year",
-       y = "Birth Rate",
-       color = "Race") +
-  scale_color_brewer(palette = "Set1", labels = c("Black", "Hispanic", "White")) +
-  theme(legend.position = "bottom")
-
-fig22<-birthhist %>%
-  mutate(whiteall = white20*10000/sumpop, blackall= black20*10000/sumpop, hispall = hisp20*10000/sumpop) %>% 
-  select(year, whiteall, blackall, hispall) %>%
-  pivot_longer(cols = c("whiteall", "blackall", "hispall"),
-               names_to = "racegroup",
-               values_to = "value"
-  ) %>%
-  ggplot(mapping = aes(x = year, y = value, color = racegroup)) +
-  geom_line() +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(2000,2020, by=4))+
-  labs(title="Age Group 20 to 24",
-       x = "Year",
-       y = "Birth Rate",
-       color = "Race",
-       caption = "Source: Birth Rates collected from
-                  CDC Vital Statistics Births Reports
-                  for 2015, 2019 and 2020.") +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "none")
-
-grid.arrange(fig21,fig22,ncol=2,nrow=1)
-
-```
-
 @fig-2 shows the trend in birth rates over the period 2000-2019 in different races for two different age groups.
@@
 On the other hand in age group 20-24 the white race group showed significanat birth rate drop from 86 to 49 birth per 1000 population., where as other race groups had much less birthrate drop.
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-4
-#| fig-cap: Changes in Other Economic factors and Birth Rate over the periods of 2004-2008 and 2014-2019
-
-#04~19 birthrate data
-birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
-pop1 = read_dta(here::here("inputs/data/pop/age_race_comp_seer.dta"))
-birth1 <- birth1 %>% 
-  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019))
-
-birth1 <- birth1 %>% 
-  subset(select=c(stname,year,numbirth1544))
-
-pop1<-pop1 %>% 
-  mutate(pop1544 = pop1519 + pop2034 + pop3544)%>% 
-  filter((year>=2004 & year<=2008) | (year>=2015 & year<=2019)) %>%
-  group_by(year,stname) %>%
-  select(pop1544) %>% 
-  summarise(across(everything(),sum))
-
-fig4_data <- right_join(birth1, pop1, by=c("stname","year"))
-fig4_data <- fig4_data %>% 
-  mutate(year = case_when(year=2004 & year<=2008 ~ 2004,
-                           year>=2015 & year<=2019 ~ 2019)) %>%
-  group_by(year,stname) %>%
-  select(numbirth1544,pop1544) %>% 
-  summarise(across(everything(),sum)) %>% 
-  mutate(birthrate=numbirth1544/pop1544*1000) %>% 
-  select(-c(numbirth1544,pop1544))
-
-fig4_data <- fig4_data %>% 
-  group_by(stname) %>% 
-  mutate(birthrate = birthrate[year == 2019] - birthrate[year == 2004])
-
-#04~19 childcare cost for fig4b
-childcare_aesc=read_dta(here::here("inputs/data/long_term/childcare_asec_09_1519.dta"))
-childcare_aesc <- childcare_aesc %>% 
-  mutate(childcare_cost=child_asec)
-childcare_aesc <- childcare_aesc %>% 
-  select(year,stname,childcare_cost)
-fig4_data <- right_join(fig4_data, childcare_aesc, by=c("stname","year"))
-fig4_data <- fig4_data %>% 
-  group_by(stname) %>% 
-  mutate(childcare_cost = childcare_cost[year == 2019] - childcare_cost[year == 2004])
-
-#04~19 rentcost data for fig4c
-rentcost_0419=read_dta(here::here("inputs/data/long_term/longdiff-rentcost-0419.dta"))
-
-rentcost_0419 <- rentcost_0419 %>% 
-  mutate(stname = fips(rentcost_0419$statefip, to='Abbreviation')) %>%
-  group_by(stname) %>% 
-  mutate(avg_grossrent = avg_grossrent[year == 2019] - avg_grossrent[year == 2004]) %>% 
-  select(year,stname,avg_grossrent)
-
-fig4_data <- right_join(fig4_data, rentcost_0419, by=c("stname","year"))
-
-#04~19 wage ratio data for fig4d
-rhs_1519=read_dta(here::here("inputs/data/long_term/longdiff-RHS-1519.dta"))
-rhs_0408=read_dta(here::here("inputs/data/long_term/longdiff-RHS-0408.dta"))
-rhs<-rbind(rhs_1519, rhs_0408)
-rhs <- rhs %>% 
-  mutate(stname = fips(rhs$statefip, to='Abbreviation')) %>% 
-  mutate(fem_male_ratio=p50wage_all/p50wage_male_all) %>% 
-  select(year,stname,fem_male_ratio) %>% 
-  group_by(stname) %>% 
-  mutate(fem_male_ratio = fem_male_ratio[year == 2019] - fem_male_ratio[year == 2004])
-
-fig4_data <- right_join(fig4_data, rhs, by=c("stname","year"))
-
-#04~19 student loan data for fig4e
-std_loan=read_dta(here::here("inputs/data/long_term/nyfed-student-debt-per-capita-0419.dta"))
-
-std_loan<- std_loan %>% 
-  group_by(stname) %>% 
-  mutate(stud_loan_cap = stud_loan_cap[year == 2019] - stud_loan_cap[year == 2004])
-
-fig4_data <- right_join(fig4_data, std_loan, by=c("stname","year"))
-
-fig4_data <- fig4_data %>% 
-  head(51) %>% 
-  select(-c(year))
-
-#drawing
-fig4b<-fig4_data %>% 
-  ggplot(aes(x=childcare_cost*1000, y=birthrate)) +
-  geom_point(size=0)+
-  labs(
-    title = "Childcare Expenditures",
-    x = "US dollar",
-    y = "Birthrate") +
-  geom_text(aes(label = stname),size=2) +
-  scale_x_continuous(limits=c(0, 5000), breaks=seq(0, 5000, by = 1000),labels = paste0("$",seq(0, 5000, by=1000))) +
-  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
-  geom_smooth(method="lm",se=FALSE, size = 0)  +
-  theme_light()
-
-fig4c <-fig4_data %>% 
-  ggplot(aes(x=avg_grossrent, y=birthrate)) +
-  geom_point(size=0)  +
-  labs(
-    title = "Rental Housing Costs",
-    x = "US dollar",
-    y = "Birthrate") +
-  geom_text(aes(label = stname),size=2) +
-  scale_x_continuous(limits=c(0, 400), breaks=seq(0, 400, by = 75),labels = paste0("$",seq(0, 400, by = 75))) +
-  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
-  geom_smooth(method="lm",se=FALSE, size = 0)  +
-  theme_light()
-
-fig4d<-fig4_data %>% 
-  ggplot(aes(x=fem_male_ratio, y=birthrate)) +
-  geom_point(size=0)  +
-  labs(
-    title = "Female-Male Wage Ratio",
-    x = "Fem to Male Wage Ratio",
-    y = "Birthrate")+
-  geom_text(aes(label = stname),size=2) +
-  scale_x_continuous(limits=c(-0.08, 0.12), breaks=seq(-0.08, 0.12, by = 0.04)) +
-  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
-  geom_smooth(method="lm",se=FALSE, size = 0)  +
-  theme_light()
-
-fig4e<-fig4_data %>% 
-  ggplot(aes(x=stud_loan_cap*1000, y=birthrate)) +
-  geom_point(size=0)  +
-  labs(
-    title = "Student Loan Debt",
-    x = "US dollar",
-    y = "Birthrate") +
-  geom_text(aes(label = stname),size=2) +
-  scale_x_continuous(limits=c(1000, 4500), breaks=seq(1000, 4500, by = 750),labels = paste0("$",seq(1000, 4500, by=750))) +
-  scale_y_continuous(limits = c(-20,10), breaks = seq(-20,10, by=5)) +
-  geom_smooth(method="lm",se=FALSE, size = 0)  +
-  theme_light()
-
-grid.arrange(fig4b,fig4c,fig4d,fig4e,ncol=2,nrow=2)
-
-```
 
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-5
-#| fig-cap: Trends in Birth Rates by Age Groups from 1990-2019
-
-# Reading in the data.
-
-birth1 = read_dta(here::here("inputs/data/nchs/nchs_births_pop_1990_2019.dta"))
-
-birth1 <- birth1 %>%
-  group_by(year) %>%
-  select(-c(stname)) %>% 
-  summarise(across(everything(),sum))
-
-birth1 <- birth1 %>% 
-  mutate(birthrate1519 = numbirth1519/pop1519*1000) %>% 
-  mutate(birthrate2024 = numbirth2024/pop2024*1000) %>% 
-  mutate(birthrate2529 = numbirth2529/(pop25+pop26+pop27+pop28+pop29)*1000) %>% 
-  mutate(birthrate3034 = numbirth3034/(pop30+pop31+pop32+pop33+pop34)*1000) %>% 
-  mutate(birthrate3539 = numbirth3539/(pop35+pop36+pop37+pop38+pop39)*1000) %>% 
-  mutate(birthrate4044 = numbirth4044/(pop40+pop41+pop42+pop43+pop44)*1000)
-
-birth1 <- birth1 %>% 
-  select(year, birthrate1519, birthrate2024, birthrate2529, birthrate3034, birthrate3539, birthrate4044) %>%
-  pivot_longer(-c(year),
-               names_to = "agegroup",
-               values_to = "birthrate"
-  )
-
-birth1<-birth1 %>% 
-  mutate(agegroup = as.character(str_replace(agegroup, "birthrate", "")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "1519", "15~19")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2024", "20~24")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "2529", "25~29")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3034", "30~34")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "3539", "35~39")))
-birth1<-birth1 %>%  mutate(agegroup=as.character(str_replace(agegroup, "4044", "40~44")))
-
-birth1 %>% 
-ggplot(mapping = aes(x = year, y = birthrate, color = agegroup)) + 
-  geom_line()  +
-  labs(
-    title = "",
-    x = "Year",
-    y = "birthrate",
-    caption = "Source: CDC Vital Statistics Births Reports.",
-    color = "Age Group"
-  ) +
-  scale_x_continuous(limits=c(1991, 2020), breaks=seq(1991, 2020, by = 8)) +
-  scale_y_continuous(limits = c(0,125), breaks = seq(0,125, by=20)) +
-  theme_light()
-```
-
 @fig-5 shows lots of important information about when people tend to have kids at certain age over the timeframe from 1990 to 2020.
@@
 
-Figure 5. Children Ever Born by Mother's Age and Birth Cohort (@fig-6).
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-6
-#| fig-cap: Children Ever Born by Mother's Age and Birth Cohort
-
-nch_cohot_data <- read_dta(here::here("inputs/data/nchs/nchs_cohort_analysis.dta"))
-agecomp_data <- read_dta(here::here("inputs/data/pop/agecomp-seer.dta"))
-
-pop_by_cohort<- agecomp_data %>%
-  select(1:32) %>%
-  group_by(year) %>%
-  select(-c(stname)) %>%
-  summarise(across(everything(),sum)) %>%
-  pivot_longer(-c(year), names_to = "mage", values_to = "pop") %>%
-  mutate(mage = as.numeric(str_replace(mage, "fem", ""))) %>%
-  mutate(birthyear = year - mage) %>%
-  mutate(cohort2 = case_when(birthyear >= 1968 & birthyear <=1972 ~ 1,
-                            birthyear >= 1973 & birthyear <=1977 ~ 2,
-                            birthyear >= 1978 & birthyear <=1982 ~ 3,
-                            birthyear >= 1983 & birthyear <=1987 ~ 4,
-                            birthyear >= 1988 & birthyear <=1992 ~ 5,
-                            birthyear >= 1993 & birthyear <=1997 ~ 6)) %>% 
-  group_by(cohort2,mage) %>%
-  summarise(pop = sum(pop))
-
-fig5_data <- right_join(nch_cohot_data, pop_by_cohort, by=c("mage", "cohort2"))
-fig5_data <- fig5_data %>%
-  mutate(birthrate=numbirth/pop*1000,
-         age_2024_year = case_when(cohort2 == 1 ~ 1992,
-                                   cohort2 == 2 ~ 1997,
-                                   cohort2 == 3 ~ 2002,
-                                   cohort2 == 4 ~ 2007,
-                                   cohort2 == 5 ~ 2012,
-                                   cohort2 == 6 ~ 2017))
-
-fig5_data <- fig5_data %>% 
-  group_by(age_2024_year) %>% 
-  mutate(cum_birthrate = cumsum(birthrate)/1000, 
-         cohort = case_when(cohort2 == 1 ~ "1968-1972        1992",
-                                  cohort2 == 2 ~ "1973-1977        1997",
-                                  cohort2 == 3 ~ "1978-1982        2002",
-                                  cohort2 == 4 ~ "1983-1987        2007",
-                                  cohort2 == 5 ~ "1988-1992        2012",
-                                  cohort2 == 6 ~ "1993-1997        2017"))
-
-fig5_data <-  na.omit(fig5_data)
-fig5_data %>%
-  ggplot(aes(x = mage,
-             y = cum_birthrate,
-             color = cohort)) +
-  geom_line()  +
-  labs(
-    title = "",
-    x = "Ages",
-    y = "Cumulative birthrate",
-    color = "Birth cohorts   Age 20 in"
-  ) +
-  scale_x_continuous(limits=c(15, 44), breaks=seq(15, 44, by = 3)) +
-  theme_minimal()+
-  theme(legend.position = "right")
-
-```
-
 @fig-6 shows lots of useful information about the cumulative birth rate by mother's age in several birth cohorts from early 1968 to late 1997.
@@
 \newpage
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-1
-#| tbl-cap: The Eight Demographic Groups that Contributed the Most to the 2007-2019 Decline in the U.S. Brith Rate
-numbirth_2044 <- read.csv(here::here("inputs/data/edu/numbirths_educ_2044.csv"))
-numbirth_2044 %>% 
-  kable(col.name=c("Age Group","Change in Birth Rate"),
-        booktabs = TRUE,show_col_types=FALSE,linesep="",
-        title = "The Eight Demographic Groups that Contributed to the Most to the 2007-2019 Decline in the U.S. Brith Rate"
-        )
-
-```
 

### 2023-03-15 outputs/paper/paper.qmd
upload graph

@@
 #Load libraries
-library(gridExtra)
-library(cdlTools)
-library(RGraphics)
+library(dplyr)
 library(tidyverse)
-library(palmerpenguins)
-library(knitr)
-library(ggplot2)
-library(haven)
-library(dplyr)
+library(readxl)
 options(dplyr.summarise.inform = FALSE)
@@
 
+\newpage
+
 ## Data Visualization
 
-Most of the figures show lots of visualization of the data.
-We can easily expect that the birth rate is declining over time within the United States due to many other factors impacting people's decision to have more kids.
-When it comes to race, there are a lot of data showing some useful information to visualize.
-Overall, black people tend to have lower birth rates than Hispanic and white people.
-Hispanics have the highest birth rates, while people are the second highest.
-But all races are showing a down trendline that the birth rates are going down over time which can be coming from other factors.
-When it comes to factors like childcare expenses, it shows that people with lower expense tends to have low birth rates, while more expense tends to have more birth rates.
-When the rental housing costs are high, the number of birth rates tends to go down, and the same for the situation in student loan debt.
-When it comes to the figures of different races, we decided to focus on age groups such as 15-19 and 20-24 because we believe that this age category has more useful information since we know that in the past, people tended to have kids much earlier than before due to having marriage at an early age.
-The birth rates by different races in the age group 15-19 have drastically decreased over time when compared to the age group 20-24.
-This is because fewer people are getting married before age 20 and having kids before age 20.
+```{r setup}
+#| echo: false
+#| message: false
+#| label: fig-1
+#| fig-cap: Work Stress by Job Status
+#Load libraries
+raw_data <- read_excel(here::here("inputs/data/GSS.xlsx"))
+clean_data <- raw_data %>% 
+  select(wrkstat, stress) %>% 
+  filter(str_detect(stress, "Always|Often|Sometimes|Hardly ever|Never"),
+         str_detect(wrkstat, "Working full time|Working part time"))
+
+positions <- c("Never", "Hardly ever", "Sometimes", "Often", "Always")
+
+clean_data %>% 
+  ggplot(aes(group = wrkstat)) +
+  geom_bar(aes(x=stress, y = ..prop.., fill=factor(..x..)))+
+  labs(y = "Percent", fill="Work Stress", x = 'Work Stress', title = "Work Stress by status" ) +
+  facet_wrap(~wrkstat) +
+  scale_x_discrete(limits = positions) +
+  scale_y_continuous(labels = scales::percent) +
+  theme_minimal(base_size = 8) +
+  theme(legend.position="none")
+```
+
+\newpage
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-2
+#| fig-cap: Work Stress by Work Hours
+raw_data <- read_excel(here::here("inputs/data/GSS.xlsx"))
+clean_data <- raw_data %>% 
+  select(hrs1, stress) %>% 
+  filter(!str_detect(hrs1, ".i|.y|.n|.d"), str_detect(stress, "Always|Often|Sometimes|Hardly ever|Never"))
+
+clean_data <- clean_data %>% 
+  mutate(range = case_when(
+    hrs1 < 20 ~ '< 20',
+    hrs1 >= 20 & hrs1 < 30 ~ '20-29',
+    hrs1 >= 30 & hrs1 < 40 ~ '30-39',
+    hrs1 >= 40 & hrs1 < 50 ~ '40-49',
+    hrs1 >= 50 & hrs1 < 60 ~ '50-59',
+    hrs1 >= 60 ~ '60 over'
+  ))
+
+positions <- c("Never", "Hardly ever", "Sometimes", "Often", "Always")
+
+clean_data %>% 
+  ggplot() +
+  geom_bar(aes(x = stress,  y = ..prop.., group = range, fill=factor(..x..))) +
+  scale_x_discrete(limits = positions) +
+  scale_y_continuous(labels = scales::percent) +
+  facet_wrap(~range) +
+  theme_minimal(base_size = 6) +
+  theme(legend.position="none")
+```
+
+\newpage
+
+```{r}
+#| echo: false
+#| label: fig-3
+#| fig-cap: Work Stress by Industry Sector
+raw_data <- read_excel(here::here("inputs/data/GSS.xlsx"))
+clean_data <- raw_data %>% 
+  select(wrkgovt, stress) %>% 
+  filter(str_detect(wrkgovt, "Government|Private"),
+         str_detect(stress, "Always|Often|Sometimes|Hardly ever|Never"))
+
+positions <- c("Never", "Hardly ever", "Sometimes", "Often", "Always")
+
+clean_data %>% 
+  ggplot() +
+  geom_bar(aes(x = stress,  y = ..prop.., group = wrkgovt, fill=factor(..x..))) +
+  scale_x_discrete(limits = positions) +
+  scale_y_continuous(labels = scales::percent) +
+  facet_wrap(~wrkgovt) +
+  theme_minimal(base_size = 8) +
+  theme(legend.position="none")
+```
+
+\newpage
+
+```{r}
+#| echo: false
+#| label: fig-4
+#| fig-cap: Work Stress by Job Satisfaction
+raw_data <- read_excel(here::here("inputs/data/GSS.xlsx"))
+clean_data <- raw_data %>% 
+  select(satjob, stress) %>% 
+  filter(str_detect(satjob, "Very satisfied|Moderately satisfied|A little dissatisfied|Very dissatisfied"),
+         str_detect(stress, "Always|Often|Sometimes|Hardly ever|Never"))
+
+positions <- c("Never", "Hardly ever", "Sometimes", "Often", "Always")
+
+clean_data %>% 
+  ggplot() +
+  geom_bar(aes(x = stress,  y = ..prop.., group = satjob, fill=factor(..x..))) +
+  scale_x_discrete(limits = positions) +
+  scale_y_continuous(labels = scales::percent) +
+  facet_wrap(~satjob) +
+  theme_minimal(base_size = 6) +
+  theme(legend.position="none")
+```
+
+\newpage
 

### 2023-03-15 scripts/Stress by Satisfication.R

@@ -17,5 +17,5 @@ clean_data %>%
   scale_x_discrete(limits = positions) +
   scale_y_continuous(labels = scales::percent) +
   facet_wrap(~satjob) +
-  theme_minimal(base_size = 12) +
+  theme_minimal(base_size = 6) +
   theme(legend.position="none")

### 2023-03-15 scripts/Stress by Status.R

@@ -13,9 +13,9 @@ positions <- c("Never", "Hardly ever", "Sometimes", "Often", "Always")
 clean_data %>% 
   ggplot(aes(group = wrkstat)) +
   geom_bar(aes(x=stress, y = ..prop.., fill=factor(..x..)))+
-  labs(y = "Percent", fill="Work Stress", x = 'Work Stress', title = "Work Stress by Work status" ) +
+  labs(y = "Percent", fill="Work Stress", x = 'Work Stress', title = "Work Stress by status" ) +
   facet_wrap(~wrkstat) +
   scale_x_discrete(limits = positions) +
   scale_y_continuous(labels = scales::percent) +
-  theme_minimal(base_size = 12) +
+  theme_minimal(base_size = 8) +
   theme(legend.position="none")

### 2023-03-15 scripts/Stress by sector.R

@@ -17,5 +17,5 @@ clean_data %>%
   scale_x_discrete(limits = positions) +
   scale_y_continuous(labels = scales::percent) +
   facet_wrap(~wrkgovt) +
-  theme_minimal(base_size = 12) +
+  theme_minimal(base_size = 8) +
   theme(legend.position="none")

### 2023-03-15 scripts/Stress by work Time.R

@@ -25,5 +25,5 @@ clean_data %>%
   scale_x_discrete(limits = positions) +
   scale_y_continuous(labels = scales::percent) +
   facet_wrap(~range) +
-  theme_minimal(base_size = 12) +
+  theme_minimal(base_size = 6) +
   theme(legend.position="none")

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include introduction

@@
 
-To measure birth rates, the National Center for Health Statistics (NCHS) aggregates data on birth rates for women by age, race, education level, and other factors; the aggregated data for a particular year is the total number of births per 1000 women.
-Each year NCHS measures birth rates nationwide and analyzes the data to make insights into societal phenomena.
-For instance, in 1997, NCHS published a paper on how a mother's education level influences the birth rate.
-In the paper, NCHS asserted that there is a negative correlation between a mother's education level and birth rate.
-Their study found a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment @citeNCHS.
+Work is the most common cause of many people to get stress.
+Work stress can be found in many forms for people.
+In this paper, we will examine job status, work hours, industry sector, satisfaction rate of your job into the correlation of work stress levels.
+There are many analysis to talk about relevant to work stress.
 
-In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S.
-population data.
-With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR].
+In this paper, we gathered data from (Publisher Name), (Data Name).
+With the gathered data, we generated figures to observe the trend in work stress of different categories using R [@citeR].
 We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages.
@@
 
-@fig-1 depicts U.S.
-birth rate trends over the last 40 years.
-Here, we can see that the trend is fluctuating relatively stable.
-However, during the post-2007 recession, it takes a steep decline with no sign of returning.
-We start our journey to explore potential explanations for the decline.
-
-In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic.
-As shown in @fig-2, by dissecting the demographics into age groups and ethnicity groups, we notice a significant decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.
-
-@sec-second-point, briefly discusses the empirical relationship between birth rates and economics and policy factors.
-We then argue that other economic, policy or social factors that have changed since 2007 do not contribute much to the decline.
-
-@sec-third-point, presents a potential attributing factor for the declining U.S.
-birth rate after the 2007 recession.
-We show strong evidence that suggests the most recent group of young adults is a significant driver of the post-2007 decline in the birth rate.
-This significance suggests that the societal shifts among young women moving through childbearing age are an ongoing trend.
-
-We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level.
-Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.
+In section 2 we discuss the source of the data used in this paper, the strengths and weaknesses the source and methodology contains, and potential blind spots that the data misses.
+Section 3 discusses the results that is being produced from the figure and what it tells us by its visualization.
+Section 4, we would provide a brief discussion about what we get from all figures, and make some thoughts.
+In section 5 there is an optional survey used to gather more information about stress caused by job status, work hours, industry sector, satisfaction rate of your job into the correlation of work stress levels.
 

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include Work Stress Levels under Discussion

@@
 
-## Birth Rates by Demographic Groups {#sec-first-point}
+## Work Stress Levels
 
-In 1997, NCHS released a report titled "Mother's Educational Level Influences Birth Rate." According to the report, a women's education level is the best predictor of how many children she will have.
-In the report, NCHS describes a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment [@citeNCHS].\
-We are particularly interested in ethnic groups of ages 15-19 and 20-24.
-@fig-2 shows a sharp decline in birth rates among all groups of women between the age of 15 to 19 during the post-2007 recession.
-@fig-2 also shows a sharp decline over the same period; however, only among white ethnic women between the ages 20 to 24, while other groups show a relatively steady decline.
-Figures that are omitted from this paper of the other age groups (25 to 34 and 35 to 44) show a similar trend both in pre and post-recession.
-However, as shown in @tbl-1, the most significant contributor to the decline in the birth rate is Hispanic, with less than High School education, women between ages 20-24 after 2007.
-This result contradicts the earlier claim made by NCHS.
-The finding suggests, in the post-2007 recession, the women's education level and birth rate now correlate positively.
+Even though work stress have different causes.
+We believe that the major cause or initial step into work stress would traffic delays.
+There are always traffic delays that most commuters would face almost every day.
+Most of these commuters are people going for work especially during rush hour time.
+Satisfaction rate is very important for jobs, and most people are sometimes not satisfy, its mainly due to these traffic delays.
+For example, if you are taking Greater Toronto Area commuters, most jobs are found in downtown Toronto, North York, and Mississauga.
+The GTA is going far west as Hamilton to east as Durham Region.
+Most of these commuters will be spending at least 1 hour of commuting to their workplace either by car or public transit.
+In an anonymized therapy session in 2014, a client expressed: “by the the time I get home from work, I’m just, you know, after that drive, I’m just so dead, it’s like, you know, shaking and don’t want to drive anymore.
+I don’t want to do anything else.
+.
+.
+sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.”
 

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include Work Stress Level in Discussion

@@
 .
-sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.”
+sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.” Traffic is the main issue that we believe that is is impacting the job satisfaction rate, which is why lots of people are facing work stress.
+We need to find a solution that can reduce the traffic, increase job satisfaction, and if possible these can expect to reduce the work stress for everyone.
 

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include Work Hours of a Work Shift & Weekly Schedule under Discussion

@@
 
-## Economic and Policy Factors {#sec-second-point}
+## Work Hours of a Work Shift & Weekly Schedule
 
-In @fig-4, we show other economic factors that might have negatively affected the birth rate in the post-2007 recession across the states.
-Although the graphs show their inverse relationship, no dramatic change is shown.
-For instance, as there were somewhere around \$1000 to \$2000 increases in childcare expenses over the past 13 years, we see that the decrease in the actual birth rate is only between 0 to -10.
-Change in per capita student loans also has increased between \$2500 to \$3250, but it only contributed to the decrease somewhere between 0 to -5 across the states.
-Based on these findings, it is hard to find an explanation of the dramatic decline in the birth rate since 2007 from economic factors.
+Right now most professionalists are working 40 hours a week, 8 hours shift each day.
+There is a plan of increasing work hours to 10 hours shift, and work only 4 days each week.
+But, based on some data, it is definitely not recommended to increase work hours because currently lots of people are always facing work stress when the number of hours are higher.
+So, if the work shift gets extended longer to reduce the total number of work days, this may cause more work stress because lots of people will be stress over number of tasks to complete during work.
+Other alternatives that can be used to reduce work stress can be reducing the total number of hours per week from 40 hours to 35 hours.
+This will allow people to face less work stess because it will reduce the number of hours for them working, and we believe 7 hours work shift can be enough to complete important tasks in a job.
 

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include new additions into Introduction

@@
 In this paper, we will examine job status, work hours, industry sector, satisfaction rate of your job into the correlation of work stress levels.
-There are many analysis to talk about relevant to work stress.
+There will be context that will motivate such as the data that is representing the factors of work stress and how people feel based on the real dataset.
+There are many analysis to talk about relevant to work stress, that can later conclude the next steps that workplace should consider to reduce work stress for many employees.
+All of the data are being collected through GSS, which allowed us to create different figures of job status, work hours, industry sector, and satisfaction rates.
+They were all written as code within the Quarto document.
+We saw some data that shows how the work stress occurs from the these factors.
+All of these information within the paper is very important because it will let the readers and others in the public to be aware of how people are facing work stress, and the analysis can provide some strong points that can lead everyone to know what employers and managers can do to avoid work stress for all employees working.
 

### 2023-03-15 outputs/paper/paper.qmd
Update paper to include new changes of Abstract

@@
     abstract: "This paper discusses data from some survey, which can gather data that is linked to how people face work stress.
-This paper uses the original data from the survey which determines the relationships between respondent work stress levels by satisfaction, status, sectors, and work time.
+It uses uses the original data from the survey which determines the relationships between respondent work stress levels by satisfaction, status, sectors, and work time.
+There will be some strong connections between the analysis and figures of the work stress.
 This paper will provide a strong analysis about the work stress that people are facing."

### 2023-03-15 outputs/paper/paper.qmd
commit before pull

--- 2023-03-15 outputs/paper/paper.qmd
@@
 
-Work is the most common cause of many people to get stress.
-Work stress can be found in many forms for people.
-In this paper, we will examine job status, work hours, industry sector, satisfaction rate of your job into the correlation of work stress levels.
-There are many analysis to talk about relevant to work stress.
+To measure birth rates, the National Center for Health Statistics (NCHS) aggregates data on birth rates for women by age, race, education level, and other factors; the aggregated data for a particular year is the total number of births per 1000 women.
+Each year NCHS measures birth rates nationwide and analyzes the data to make insights into societal phenomena.
+For instance, in 1997, NCHS published a paper on how a mother's education level influences the birth rate.
+In the paper, NCHS asserted that there is a negative correlation between a mother's education level and birth rate.
+Their study found a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment @citeNCHS.
 
-In this paper, we gathered data from (Publisher Name), (Data Name).
-With the gathered data, we generated figures to observe the trend in work stress of different categories using R [@citeR].
+In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S.
+population data.
+With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR].
 We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages.
@@
 
-In section 2 we discuss the source of the data used in this paper, the strengths and weaknesses the source and methodology contains, and potential blind spots that the data misses.
-Section 3 discusses the results that is being produced from the figure and what it tells us by its visualization.
-Section 4, we would provide a brief discussion about what we get from all figures, and make some thoughts.
-In section 5 there is an optional survey used to gather more information about stress caused by job status, work hours, industry sector, satisfaction rate of your job into the correlation of work stress levels.
+@fig-1 depicts U.S.
+birth rate trends over the last 40 years.
+Here, we can see that the trend is fluctuating relatively stable.
+However, during the post-2007 recession, it takes a steep decline with no sign of returning.
+We start our journey to explore potential explanations for the decline.
+
+In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic.
+As shown in @fig-2, by dissecting the demographics into age groups and ethnicity groups, we notice a significant decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.
+
+@sec-second-point, briefly discusses the empirical relationship between birth rates and economics and policy factors.
+We then argue that other economic, policy or social factors that have changed since 2007 do not contribute much to the decline.
+
+@sec-third-point, presents a potential attributing factor for the declining U.S.
+birth rate after the 2007 recession.
+We show strong evidence that suggests the most recent group of young adults is a significant driver of the post-2007 decline in the birth rate.
+This significance suggests that the societal shifts among young women moving through childbearing age are an ongoing trend.
+
+We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level.
+Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.
 
@@
 #| echo: false
+#| warning: false
 #| message: false
@@
 #| echo: false
+#| warning: false
 #| label: fig-3
@@
 #| echo: false
+#| warning: false
 #| label: fig-4
@@
 
-## Work Stress Levels
+## Birth Rates by Demographic Groups {#sec-first-point}
 
-Even though work stress have different causes.
-We believe that the major cause or initial step into work stress would traffic delays.
-There are always traffic delays that most commuters would face almost every day.
-Most of these commuters are people going for work especially during rush hour time.
-Satisfaction rate is very important for jobs, and most people are sometimes not satisfy, its mainly due to these traffic delays.
-For example, if you are taking Greater Toronto Area commuters, most jobs are found in downtown Toronto, North York, and Mississauga.
-The GTA is going far west as Hamilton to east as Durham Region.
-Most of these commuters will be spending at least 1 hour of commuting to their workplace either by car or public transit.
-In an anonymized therapy session in 2014, a client expressed: “by the the time I get home from work, I’m just, you know, after that drive, I’m just so dead, it’s like, you know, shaking and don’t want to drive anymore.
-I don’t want to do anything else.
-.
-.
-sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.” Traffic is the main issue that we believe that is is impacting the job satisfaction rate, which is why lots of people are facing work stress.
-We need to find a solution that can reduce the traffic, increase job satisfaction, and if possible these can expect to reduce the work stress for everyone.
+In 1997, NCHS released a report titled "Mother's Educational Level Influences Birth Rate." According to the report, a women's education level is the best predictor of how many children she will have.
+In the report, NCHS describes a direct relationship between years of education and birth rates, with the highest birth rates among women with the lowest educational attainment [@citeNCHS].\
+We are particularly interested in ethnic groups of ages 15-19 and 20-24.
+@fig-2 shows a sharp decline in birth rates among all groups of women between the age of 15 to 19 during the post-2007 recession.
+@fig-2 also shows a sharp decline over the same period; however, only among white ethnic women between the ages 20 to 24, while other groups show a relatively steady decline.
+Figures that are omitted from this paper of the other age groups (25 to 34 and 35 to 44) show a similar trend both in pre and post-recession.
+However, as shown in @tbl-1, the most significant contributor to the decline in the birth rate is Hispanic, with less than High School education, women between ages 20-24 after 2007.
+This result contradicts the earlier claim made by NCHS.
+The finding suggests, in the post-2007 recession, the women's education level and birth rate now correlate positively.
 
-## Work Hours of a Work Shift & Weekly Schedule
+## Economic and Policy Factors {#sec-second-point}
 
-Right now most professionalists are working 40 hours a week, 8 hours shift each day.
-There is a plan of increasing work hours to 10 hours shift, and work only 4 days each week.
-But, based on some data, it is definitely not recommended to increase work hours because currently lots of people are always facing work stress when the number of hours are higher.
-So, if the work shift gets extended longer to reduce the total number of work days, this may cause more work stress because lots of people will be stress over number of tasks to complete during work.
-Other alternatives that can be used to reduce work stress can be reducing the total number of hours per week from 40 hours to 35 hours.
-This will allow people to face less work stess because it will reduce the number of hours for them working, and we believe 7 hours work shift can be enough to complete important tasks in a job.
+In @fig-4, we show other economic factors that might have negatively affected the birth rate in the post-2007 recession across the states.
+Although the graphs show their inverse relationship, no dramatic change is shown.
+For instance, as there were somewhere around \$1000 to \$2000 increases in childcare expenses over the past 13 years, we see that the decrease in the actual birth rate is only between 0 to -10.
+Change in per capita student loans also has increased between \$2500 to \$3250, but it only contributed to the decrease somewhere between 0 to -5 across the states.
+Based on these findings, it is hard to find an explanation of the dramatic decline in the birth rate since 2007 from economic factors.
 

### 2023-03-16 outputs/paper/paper.qmd
2.1 and 2.2 done

@@
 
-The datasets for this paper were obtained from the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper], a paper from the Journal of Economic Perspectives [@citewebsite].
-The data used by the paper was compiled from various sources such as the NCHS Natality Database [@citenchs], CDC SEER [@citecdcseer], the Current Population Survey Annual Social and Economic Supplement (CPS ASEC) [@citeipums], the New York Federal Reserve [@citefed], and the Integrated Public Use Microdata Series (IPUMS) [@citeipums].
-The paper and datasets were published in the year 2022.
-These various sources provide a lot of useful information, shown throughout all of the figures we created in this paper.
-For example, the sources collected lots of raw data such as the population of people giving birth, race, lifestyle factors, age of parent when giving birth, etc.
-So, all of this information was collected from these various sources and written into the paper "The Puzzle of Falling US Birth Rates since the Great Recession" [@citepaper].
-Which was then analyzed in our paper with our version of the analysis.
+The dataset for this paper were obtained from the US General Social Survey [] from NORC at the University of Chicago.
+The dataset was published in the year 2021.
+The raw dataset includes data from 568 Survey Questions for 4032 Survey Participants.
+GSS monitors pubic opinion and behavior in the United States of America.
+This survey has been conducted since 1972 by the NORC at the University of Chicago and funded by the National Science Foundation(NSF).
+We have selected 10 items from the raw dataset, and only extracted selected variables.
+These selected variables were imported in R using the package haven[@citehaven] and readxl[@citereadxl].
+Using the R[@citeR] package tidyverse[@citetidyverse] and dplyr[@citedplyr], we cleaned and performed exploratory data analysis on the selected dataset to get insights into the data.
+The insights gained from this data set is shown throughout all of the figures we created in this paper.
+The some of the selected items are year, travel time to work, labor force status, number of hours worked, and etc.
+The analysis of these collected information is available on section 2.3 Data Visualization and Analysis.
 
-## Data Cleaning
+## Data Cleaning and Modification
 
-Most of the data used in this paper were cleaned to get the most accurate information and remove any unnecessary information to be outputted in this paper.
-Data were cleaned to remove data of some years, age groups, race categories, etc.
-So that the revised figures can provide the most accurate information according to our analysis and not have any information that is unrelevant to our analysis.
-From our dataset, we dropped any values that are not related to the specific category, for example, values like NA.
-We have tried to filter our data much as we can for all of the graphs to make sure that it can make sense.
+The data used in this paper were cleaned to get an accurate representation and remove any unnecessary information in this paper.
+Data were cleaned to remove data of some of results that does not have responds on specific items that we are looking into.
+Some of participants refused to answer or skipped some of the questions for various reasons.
+We filtered out data that would not have meaningful result on the graphs.
 
-## Data Modification
-
-We have included mainly all datasets and variables that are very important for our research question.
-We included variables by race, birth rates, age group, and other important factors.
-Lots of other variables were included, such as the total number of births and birth rates.
-Both of these allow us to analyze the decline of births over time.
-We did lots of modifications of the variables for this paper, which were mainly done within the age category.
-Generally, the original paper has an age category as one thing from age 15-44.
-For our figures, we made several figures and chose a specific age range.
-We have tried our best to divide the range into categories like 15-19, 20-24, 25-34, and 35-44 throughout most of the figures.
-We have also modified for race by having a figure that shows the data by different races, such as Black, White, and Hispanic people.
-Several figures had an important modifications to be relevant to our analysis throughout the paper.
+For each figures, we selected certain items that are closely related to work condition and explored how they affects the work stress.
+The Data set is divided into categories and range, like working 20 or less, 20-29, 30-39, 40-49, 50-59, 60 over hours of working per week, or partime/fulltime worker.
+Eachfigures explores different categorization to see how they effect personal stress of work.
 
@@
 
-## Data Visualization
+## Data Visualization and Analysis
 
@@
 ```
+
+@fig-1 shows the work stress by work status.
 
@@
 
-@fig-2 shows the trend in birth rates over the period 2000-2019 in different races for two different age groups.
-In all race group, the birth rates starts to drop significantly after the Great Recession.
-In a age group of 15-19, all three race groups had similar drops on the birth rate.
-On the other hand in age group 20-24 the white race group showed significanat birth rate drop from 86 to 49 birth per 1000 population., where as other race groups had much less birthrate drop.
-
-@fig-4 shows an interesting relationship between changes in birthrate and other economic factors over five years period. Each marker represents a state labelled with its two-letter abbreviation.
-
-Annual expenditures on childcare for families with children under 12 who report positive childcare spending rose \$2,170 nationwide for the average of 2015--2019, based on data from the Current Population Survey (all dollar values throughout are measured in constant 2019 dollars).
-But in the cross-state correlations, places where childcare expenditures increased more, did not experience a noticeable drop in birth rates.
-In fact, the trend shows that more increases in childcare expenditure have no significant impact on birthrate changes.
-
-Over this period, average monthly rents for a two- to three-bedroom apartment rose \$124 per month (from \$930 to \$1,060 in 2018 dollars, a 14 percent increase) nationwide.
-The increase was much larger in some states, like Colorado, Washington, and the District of Columbia.
-The data, however, do not indicate a negative association between rent changes and birth rate changes.
-
-We used the female-to-male median earnings ratio among full-time, year-round workers to examine women's economic opportunities.
-On average, the median earning ratio rose to 0.04 in the period.
-12 states where the wage ratio rose the most over this sample period did not show a greater birth rate decline.
-
-If young adults are saddled with debt, they might not have sufficient disposable income to have a child or more children.
-We consider the total student debt per capita level in a state, which has increased from \$2,500 to \$5,400 (in 2018 dollars) on average.
-The relationship between state-level student debt and the birth rate is generally flat.
-It indicates that increases in student debt are related to aggregate reductions in birth rates.
-
-@fig-5 shows lots of important information about when people tend to have kids at certain age over the timeframe from 1990 to 2020.
-Based on the graph, the data shows that more people have kids after the age of 30 during this present time.
-In the past, most people tend to have kids before the age of 30.
-Around 2000 to 2005, the lifestyle started to change, where people tended to either get married or have kids later, which is why the trend of birth rates in age 30 and up has been going higher since the 2000s.
-While births before the age of 30 tend to slow down and decrease over time from the 2000s.
-According to the graph, maybe in another 100 years, we can expect that most birth rates can happen around 40-44.
-
-There are drastic changes shown in @fig-5, such as age 30-44 tends to increase around the 2000s, while the ages below 30 decreases for the number of birth rates.
-Around the decade of the 2000s or the beginning of this new 21st century, there was lots of modernization from people's lifestyle to other things on Earth, which is coming from several factors such as education and employment due to new innovations and inventions [@citedimock].
-There are lots of new behaviour being introduced to a new generation of people.
-So, as they get new behaviour of attitudes, that is being involved in their decisions of marriages, and when to have kids.
-This is why lots of people are now getting married late and/or having kids much later when compared to the past.
-
-@fig-6 shows lots of useful information about the cumulative birth rate by mother's age in several birth cohorts from early 1968 to late 1997.
-We tend to use these years since, in recent times, we believe that there is enough data for these years since most of the births and living people are among these cohorts.
-According to @fig-6, the earlier the cohort is, the cumulative birth rate will be really high.
-While the later the cohort is, the cumulative birth rate will be much lower.
-The highest cumulative birth rate is coming from the birth cohort of 1968-1972 and 1973-1977.
-While the lowest comes from the 1993-1997 birth cohort.
-\newpage
-
-@tbl-1 shows the most contributor to the decline in birth rates across all age groups.
-The top contributor is Hispanic with less than high school education women between age 20-24 at -88.3, Hispanic with high school education women between age 15-19 at -57.5, and Black (Non-Hispanic) women between age 15-19 at -39.8, respectively.
-
-\newpage
-
 # Discussion

### 2023-03-16 scripts/Stress by sector.R

@@ -1,6 +1,6 @@
 library(dplyr)
 library(tidyverse) 
-
+library(readxl)
 
 raw_data <- read_excel("inputs/data/GSS.xlsx")
 clean_data <- raw_data %>% 

### 2023-03-16 outputs/paper/paper.qmd
data part done

@@
 @fig-1 shows the work stress by work status.
+To visualize this figure, work stress and work status data was selected and grouped by the response.
+With the cumulative results by responds, two separate bar graph was created to show the work stress people experienced based on their working status.
+
+@fig-1 shows that the full-time workers have tendency to experience more work stress then the part-time workers.
+The full-time workers often have more work experience therefore have more responsibility then the part-time workers.
+This would expose them to more stressful situation often compared to the part-time workers.
+Also, part-time workers have other aspects of life and often focuses more on those, for examples school or house work.
+It makes them to care less about the work and receives less stress from it.
+Last but not least, part-time workers have much more flexible work schedule then the full-time workers.
+This prevents them from burn out from the work and potentially reduces stress.
 
@@
 ```
+
+@fig-2 shows more detailed categorization based on the working hours.
+Generally, 35 hours per week is the standard to separate part-time workers and full-time workers.
+However, labeling a person who works 8 hours per week and 30 hours per week as same as part-time worker does not provides insightful results.
+Therefore we further categorized into more detailed range.
+It shows the work stress of workers based on the 6 categories of working hours.
+
+It is clearly visible that work stress has linear relationship with the working hours.
+As more hours workers work every week, they have more chance to experience stress.
+Working hours have close relationship with the burn out issues of the worker.
+More work hours leads them to less time to spend for self development and interest which is a great source of the stress.
 

### 2023-03-16 outputs/paper/paper.qmd
data part done 2

@@
 @fig-2 shows more detailed categorization based on the working hours.
+Work hour and work stress are selected to visualized this data.
 Generally, 35 hours per week is the standard to separate part-time workers and full-time workers.
@@
 It shows the work stress of workers based on the 6 categories of working hours.
+Answers that are no aplicable to this sections is cleaned to represent best result.
 
@@
 
+@fig-3 shows the work stress received for a people who works in the government sector and private sector.
+working government sector and work stress are selected for this visualization.
+Some of industries are hard to categorize into either government or non-government, therefor any answers that are not clear or unanswered were eliminated to have clear vision on this category.
+The data is grouped based on the responds on working sector.
+
+This graph show that people who works in a government sector is more likely to get stress then the people who works in a private sector.
+The key difference between the government sector and the private sector is the salary.
+Government jobs tend to have lower salary then the same level jobs in the private sector.
+However, the difference evens out with benefits like a pension and stable work condition.
+This results work stress is more affected by current status of the work and pay, rather then the future benefit or the work safety.
+
 \newpage
@@
 
-\newpage
+@fig-4 shows the relationship between job satisfaction and the work stress.
+These two items are selected and grouped by answers.
+The survey on the job satisfaction is done with 4 steps from very satisfied to very dissatisfied, and the work stress is done with 5 steps from Never to Always.
+To visualize different work stress by job satisfaction, four graphs are drawn based on the levels of satisfaction on the job.
 
-# Results {#sec-results}
-
-@fig-1 shows that over time, the number of births in the United States is dropping steadily, while it slightly went up on peak around 1990, 2001, and 2007.
-Most likely, the hikes going up during these years might have been into other factors that made more people among the selected population have kids.
-After 2007, the number of births drastically dropped a lot.
-Nowadays, among 1,000 women from age 15 to 44, only about 56 births will occur.
-While in the past, around the 1980s and 1990s, the number of births was around the 65 to 70 range.
-The line in the graph shows that in the near future, there will be a much lower amount of people having kids, and this can be due to other factors from their lifestyle, which is making them decide not to have more kids.
-
-While the figure shows the birth rate trend is going down, there are certain spikes in the peak of birth rates around 1990, 2001, and 2007.
-There were some incidents that must have happened in the United States in those years, which may have caused people to stay home more often.
-As we know that people that are likely to be unemployed tend to have more kids as they will have more freedom during their time spent at home.
-This allows the birth rates to be higher than usual.
-During these years, Americans faced lower employment opportunities within society [@citecatholic].
-They do not have access to useful resources that improve their lifestyle in these years, so they are at home most of the time, which encourages couples to have an interest in having more kids, which ends up increasing the birth rates for these years.
+For the people who either very satisfies or very dissatisfies on their job has clear differences on work stress.
+As expected, people who are very satisfied on their job gets significantly less stress then the people who are very dissatisfied.
+This trend continues on the little satisfied and moderately satisfied, but there is a finding worth to look at.
+Although people who are a little dissatisfied with their jobs have little higher chance to get more stress, but there is a little to no difference in a chance to get no stress.
+This finding shows that job satisfaction has clear relationship with work stress but if its satisfaction is not extreme in either way  people who have resistance to the work stress have little effect from it.
 

### 2023-03-16 outputs/paper/paper.qmd
Update paper

@@
   - Youngho Kim
-thanks: "Code, data are available at: https://github.com/YoungKim164/workstress and https://doi.org/10.1257/jep.36.1.151 and replication package at: https://doi.org/10.48152/ssrp-cdf6-2r92"
+thanks: "Code, and data are available at: https://github.com/YoungKim164/workstress"
 date: "`r Sys.time()`"
@@
     colorlinks: true
-    lot: true
+    lot: false
     lof: true
@@
 
-The dataset for this paper were obtained from the US General Social Survey [] from NORC at the University of Chicago.
+The dataset for this paper were obtained from the US General Social Survey [@norc] from NORC at the University of Chicago.
 The dataset was published in the year 2021.

### 2023-03-16 outputs/paper/paper.qmd
Update paper to include all parts of Dicussion

@@
 \newpage
+
+# Results
 
@@
 
-## A Potential Explanation: Shifting Priorities {#sec-third-point}
+## Conclusion
 
-When it comes to the age of the population of the United States and comparing it from the past to now.
-It shows that more people tend to have kids much later than before.
-While fewer people have kids earlier than in the past.
-In the past, more people tend to have kids much earlier.
-There are several reasons why more people had kids at an early age back then; people used to get married much earlier, and many people were unemployed or doing a job for self-employment, which was not even an educated or professional job.
-The modern lifestyle that has been adopted in the past few decades tends to make people focus on all other areas of life, which makes them marry much later, and have kids later.
-Nowadays, many marriages happen at the age of late 20s or early 30s, while they tend to have their first child around the early to mid-30s.
-@fig-5 depicts the birth rates in women between the ages 30-39 and 35-39 positive slope trend.
-Whether pre or post-2007 recession, this trend looks steady over the past 40 years.
-This finding aligns with the current common belief and the social norm where, perhaps, women's peak childbearing ages have shifted from the early twenties to early thirties.
-@fig-6 shows the average number of children born to women by age in five-year intervals from the 1968-1972 through 1993-1997 birth cohorts.
-These birth cohorts include women entering their peak childbearing years (20--24) in 1992, 1997, 2002, 2007, and 2017.
-This chart indicates that the three cohorts of young adult women born between 1968 and 1982 in 1992, 1997, and 2002 had similar childbearing age profiles.
-The 2007 birth cohort of women had fewer children in their 20s and 30s.
-The 2012 and 2017 birth cohorts (born between 1988 and 1997) have significantly fewer children than the preceding cohorts.
-This finding is consistent with a decline in births, primarily to more recent cohorts of women.
-From a statistical point of view, this divergence across cohorts explains the sizable decrease in annual birth rates that began in 2007.
-
-## Weakness and Moving Forward
-
-### Bias and Ethical Concerns
-
-As we know that this data was collected and created by other human beings, just as how we are.
-There are some ethical concerns that could affect the accuracy of our data because humans can make mistakes, and they can make decisions based on their preferences, which can affect the whole dataset.
-So, this analysis is only valid based on our dataset, which can be modified if the original dataset is changed.
-
-The original data shows different races, such as White, Black, and Hispanic.
-They have included Hispanic based on all colours, which erases the identity of Black and Indigenous Hispanic people.
-This tends to be a bit of racism because the information provided on the data did not specifically indicate whether or not the Hispanic category includes everyone that identifies as Hispanic or is it just people that look Hispanic.
-
-### Data Collection Concerns
-
-This dataset tends to collect huge data information on the population among people from the United States and considers lots of factors from race, birth rates by age, student debts, etc.
-Having too much information means that when the data is collected, there is a possibility of human error, calculation error, and chances of small modification of data due format of reporting an error.
-So, the way how the data was collected can have lots of problems because people can make mistakes easily.
-
-### Moving Forward
-
-This paper raised concerns about a dramatic decline in U.S.
-birth rates starting in 2007.
-Through the paper, we proposed a general explanation for the decline in births across recent cohorts of U.S.
-women that focuses on the shifting priorities of cohorts.
-As dictated by the recent socio-phenomena that encompasses preferences for having children, life aspirations, and the nature of parenting, a change in the relationship between education and childbearing, a rise in childlessness, and the establishment of a two-child norm for those having children have irreversibly changed [@citepaper].
-This new social phenomenon that might be causing the dramatic decline in U.S.
-birth rates is important from an economic policy standpoint to acknowledge that an aging population and shrinking workforce pose challenges to economic growth and the sustainability of social insurance systems [@Yang2003].
-Thus, the most appropriate way to address declining U.S.
-birth rates may be required: 1.
-developing A.I.
-and robots where a shrinking workforce poses a problem, 2.
-an appropriate policy to address issues with Social Security and Medicare on a secure basis for the long-term.
-Suppose the U.S.
-birth rate is not showing any reverse trend shortly.
-In that case, the U.S.
-economy and political system should enforce greater emphasis on addressing the issue before it becomes irreversible.
+Work stress seems to be the main problem for everyone in both physical and mental health.
+Through our data, there was a lots of things to imaginate about the causes of work stress which is revealed throughout the paper.
+The causes in correlation into work stress are by factors such as job status, work hours, industry sector, and satisfaction rates.
+Half of the people working either full time or part time are facing work stress, and this could be because of their assigned tasks that they have to do every day.
+Under work hours, the higher the hours people work for, the higher amount of people facing work stress.
+This could be because of people working for too long, which is affecting their health condition, and that causes them to face higher levels of work stress.
+Government sector tends to be slightly higher than private sector for work stress, but this could be because government sectors tends to pay a bit less compared to bigger private sectors.
+For job satisfaction, there are some people that are often disatisfied, which reflects their work stress and this could be because they job they are working for is not suitable for them, lack of interst, commute issues, traffic delays, work schedule, etc.
+So, there are lots of issues in work, which is the main issue of work stress.
+After looking over everything, we believe that if companies within the same job industry pays the same salary, reducing the work hours to 35 hours a week, and trying to make some jobs hybrid or remote.
+Then, this can reduce the number of work stress people are facing every day.
 

### 2023-03-16 outputs/paper/paper.qmd
Update paper and references

@@
 
+\newpage
+
 # Introduction
@@
 There are many analysis to talk about relevant to work stress, that can later conclude the next steps that workplace should consider to reduce work stress for many employees.
-All of the data are being collected through GSS, which allowed us to create different figures of job status, work hours, industry sector, and satisfaction rates.
+All of the data are being collected through the US General Social Survey [@norc] from NORC at the University of Chicago, which allowed us to create different figures of job status, work hours, industry sector, and satisfaction rates.
 They were all written as code within the Quarto document.
@@
 
-In this paper, we gathered data from (Publisher Name), (Data Name).
 With the gathered data, we generated figures to observe the trend in work stress of different categories using R [@citeR].
@@
 
-In section 2 we discuss the source of the data used in this paper, the strengths and weaknesses the source and methodology contains, and potential blind spots that the data misses.
+In Section 2 we discuss the source of the data used in this paper, the strengths and weaknesses the source and methodology contains, and potential blind spots that the data misses.
 Section 3 discusses the results that is being produced from the figure and what it tells us by its visualization.
@@
 Eachfigures explores different categorization to see how they effect personal stress of work.
-
-\newpage
 
@@
 The GTA is going far west as Hamilton to east as Durham Region.
-Most of these commuters will be spending at least 1 hour of commuting to their workplace either by car or public transit.
+Most of these commuters will be spending at least 1 hour of commuting to their workplace either by car or public transit [@trafficsatisfaction].
 In an anonymized therapy session in 2014, a client expressed: “by the the time I get home from work, I’m just, you know, after that drive, I’m just so dead, it’s like, you know, shaking and don’t want to drive anymore.
@@
 Right now most professionalists are working 40 hours a week, 8 hours shift each day.
-There is a plan of increasing work hours to 10 hours shift, and work only 4 days each week.
+There is a plan of increasing work hours to 10 hours shift, and work only 4 days each week [@workstress].
 But, based on some data, it is definitely not recommended to increase work hours because currently lots of people are always facing work stress when the number of hours are higher.
@@
 
+# Appendix
+## Appendix A: Supplementary Survey
+There are many researchers finding way to reduce work stress for everyone, and finding the major causes that is making people to face work stress.
+
+If you complete this survey, then that means you are giving us permission by collecting your basic information and understand that we will use your responses in our study.
+The survey is anonymous and will stay confidential by not sharing it with anyone else.
+We will not collect or share any personal identifiable information.
+The survey is also voluntary, and you may withdraw at any time.
+
+If you have any questions or concerns, feel free to contact Sagith Kalaichelvam (sagith [dot] kalaichelvam [at] mail [dot] utoronto [dot] ca).
+
+1.  What is your age?
+    a. 18-24
+    b. 25-34
+    c. 35-44
+    d. 45-64
+    e. 65+
+2.  What is your gender?
+    a. Male
+    b. Female
+    c. Others
+    d. Prefer not to say
+3.  Which causes do you think is the main cause of you facing work stress?
+    a. Traffic delays
+    b. Too much work tasks
+    c. Long hours
+    d. Not interested in the job
+    e. Others
+4.  How many hours do you prefer to work for the week?
+    a. 25 and under hours
+    b. 26-30 hours
+    c. 31-35 hours
+    d. 36-40 hours
+5.  How many hours would you work in a entire shift for the day?
+    a. 4-6 hours
+    b. 7-8 hours
+    c. 9-10 hours
+    d. 11-12 hours
+6.  How often do you get stress at work?
+    a. Always
+    b. Most of the times
+    c. Sometimes
+    d. Few times
+    e. Never
+7.  How long does it take for you to commute to work?
+    a. 15-20 mins
+    b. 30 mins
+    c. 1 hour 
+    d. More than 1 hour
+8.  Does your boss or manager make you to work on a lots of tasks on a single work shift?
+    a. Yes
+    b. Sometimes
+    c. Never
+    d. Not Applicable
+
+Thank you for participating in this survey! We hope that your responses will be very helpful in our research study to make some analysis.
+
 # References

### 2023-03-16 scripts/00-Simulation_Stress by Satisfication.R
Update all Simulation scripts

@@ -1,7 +1,14 @@
+#### Preamble ####
+# Purpose: Read in data from the GSS
+# Prerequisites: Need to know where to get GSS data
+# Author: Sagith Kalaichelvam and Youngho Kim
+# Email: sagith.kalaichelvam@mail.utoronto.ca
+# Date: 16 March 2023
+# GitHub: https://github.com/YoungKim164/workstress
 library(dplyr)
 library(tidyverse) 
 
-
+# Reading the file
 raw_data <- read_excel("inputs/data/GSS.xlsx")
 clean_data <- raw_data %>% 
   select(satjob, stress) %>% 

### 2023-03-16 scripts/00-Simulation_Stress by Status.R

@@ -1,7 +1,13 @@
-library(dplyr)
+#### Preamble ####
+# Purpose: Read in data from the GSS
+# Prerequisites: Need to know where to get GSS data
+# Author: Sagith Kalaichelvam and Youngho Kim
+# Email: sagith.kalaichelvam@mail.utoronto.ca
+# Date: 16 March 2023
+# GitHub: https://github.com/YoungKim164/workstresslibrary(dplyr)
 library(tidyverse) 
 
-
+# Reading the file
 raw_data <- read_excel("inputs/data/GSS.xlsx")
 clean_data <- raw_data %>% 
   select(wrkstat, stress) %>% 

### 2023-03-16 scripts/00-Simulation_Stress by sector.R

@@ -1,7 +1,15 @@
+#### Preamble ####
+# Purpose: Read in data from the GSS
+# Prerequisites: Need to know where to get GSS data
+# Author: Sagith Kalaichelvam and Youngho Kim
+# Email: sagith.kalaichelvam@mail.utoronto.ca
+# Date: 16 March 2023
+# GitHub: https://github.com/YoungKim164/workstress
 library(dplyr)
 library(tidyverse) 
 library(readxl)
 
+# Reading the file
 raw_data <- read_excel("inputs/data/GSS.xlsx")
 clean_data <- raw_data %>% 
   select(wrkgovt, stress) %>% 

### 2023-03-16 scripts/00-Simulation_Stress by work Time.R

@@ -1,7 +1,14 @@
+#### Preamble ####
+# Purpose: Read in data from the GSS
+# Prerequisites: Need to know where to get GSS data
+# Author: Sagith Kalaichelvam and Youngho Kim
+# Email: sagith.kalaichelvam@mail.utoronto.ca
+# Date: 16 March 2023
+# GitHub: https://github.com/YoungKim164/workstress
 library(dplyr)
 library(tidyverse) 
 
-
+# Reading the file
 raw_data <- read_excel("inputs/data/GSS.xlsx")
 clean_data <- raw_data %>% 
   select(hrs1, stress) %>% 

### 2023-03-16 scripts/01-test_data.R
Upload 01-test_data

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Work Stress
+# Author: Sagith Kalaichelvam and Youngho Kim
+# Data: 16 March 2023
+# Contact: sagith.kalaichelvam@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(janitor)
+library(here)
+
+#### Test data ####
+# If there is TRUE outputting, then test is assumed to be passed
+# All of the test are passing with return value of True
+# For those with multiple data outputting, there might be some false outputing
+# after the True value, those can be ignored because as long the first value
+# is outputting True, then that means it is passing the test, since we are only
+# checking for first value to be outputted as True in the data
+
+# Checking if the year matches the year listed in the dataset
+raw_data$year |> unique() == c('1972')
+
+# Test the year data is a numeric
+raw_data$year |> class() == "numeric"
+
+# Checking if the workstat info matches the workstat info listed in the dataset
+raw_data$wrkstat |> unique() == "Working full time"
+
+# Checking if the sat info of job  info matches the stat info of each job listed in the dataset
+raw_data$satjob |> unique() == "A little dissatisfied"
+
+

### 2023-03-16 scripts/01-test_data.R
Upload 01-test_data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Work Stress
 # Author: Sagith Kalaichelvam and Youngho Kim
-# Data: 16 March 2023
+# Data: 15 March 2023
 # Contact: sagith.kalaichelvam@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-03-16 scripts/01-test_data.R
Upload testing

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Work Stress
 # Author: Sagith Kalaichelvam and Youngho Kim
-# Data: 15 March 2023
+# Data: 16 March 2023
 # Contact: sagith.kalaichelvam@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-03-17 outputs/paper/paper.qmd
Update paper

@@
     geometry: margin = 1in
-    abstract: "This paper discusses data from some survey, which can gather data that is linked to how people face work stress.
+    abstract: "This paper discusses data from US General Social Survey, which can gather data that is linked to how people face work stress.
 It uses uses the original data from the survey which determines the relationships between respondent work stress levels by satisfaction, status, sectors, and work time.
@@
 Most of these commuters will be spending at least 1 hour of commuting to their workplace either by car or public transit [@trafficsatisfaction].
-In an anonymized therapy session in 2014, a client expressed: “by the the time I get home from work, I’m just, you know, after that drive, I’m just so dead, it’s like, you know, shaking and don’t want to drive anymore.
+In an anonymized therapy session in 2014, a client expressed: “by the time I get home from work, I’m just, you know, after that drive, I’m just so dead, it’s like, you know, shaking and don’t want to drive anymore.
 I don’t want to do anything else.
@@
 .
-sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.” Traffic is the main issue that we believe that is is impacting the job satisfaction rate, which is why lots of people are facing work stress.
+sometimes I like if there’s been lots of traffic or like if I’ve been tired and like been forcing myself to stay awake so I can get home.” [@therapy] Traffic is the main issue that we believe that is is impacting the job satisfaction rate, which is why lots of people are facing work stress.
 We need to find a solution that can reduce the traffic, increase job satisfaction, and if possible these can expect to reduce the work stress for everyone.
@@
 This will allow people to face less work stess because it will reduce the number of hours for them working, and we believe 7 hours work shift can be enough to complete important tasks in a job.
+
+## Weaknesses and Next Steps
+
+There are chances to have some weakness within the dataset collected.
+The data within the dataset are among a specific unknown group of people, where we can not exactly expect the response will be the same right now.
+People tend to change their decisions often, so the data in the dataset is only 100% valid at the time when it was first released.
+So, our entire analysis is being supported based on this dataset and more useful when it was intially released.
+The data is coming from 2021, but that is already 2 years from now.
+So, in the future, US General Social Survey needs update their datasets either annually or monthly to get the most accurate data, which can provide a stronger analysis and more reasonable to our paper.
 
@@
 
+\newpage
+
 # References

### 2023-03-17 outputs/paper/paper.qmd
Update paper

@@
   - Youngho Kim
-thanks: "Code, and data are available at: https://github.com/YoungKim164/workstress"
+thanks: "Code and data are available at: https://github.com/YoungKim164/workstress"
 date: "`r Sys.time()`"
@@
 Even though work stress have different causes.
-We believe that the major cause or initial step into work stress would traffic delays.
+We believe that the major cause or initial step into work stress would be traffic delays, and that affects the job satisfaction rate.
 There are always traffic delays that most commuters would face almost every day.

### 2023-03-17 outputs/paper/paper.qmd
Update paper to include changes on survey

@@
 ## Appendix A: Supplementary Survey
+
+Supplementary Survey is on Google Forms: https://forms.gle/fdPuTeWRb7M68Z8ZA
+
+### Survey Introduction
+
 There are many researchers finding way to reduce work stress for everyone, and finding the major causes that is making people to face work stress.
@@
 If you have any questions or concerns, feel free to contact Sagith Kalaichelvam (sagith [dot] kalaichelvam [at] mail [dot] utoronto [dot] ca).
+
+### Survey Questions
 
@@
     b. Female
-    c. Others
-    d. Prefer not to say
+    c. Prefer not to say
+    d. Other...
 3.  Which causes do you think is the main cause of you facing work stress?
@@
     d. Not interested in the job
-    e. Others
+    e. Other...
 4.  How many hours do you prefer to work for the week?
@@
     d. 36-40 hours
+    e. Other...
 5.  How many hours would you work in a entire shift for the day?
@@
 
+### Survey Conclusion
+
 Thank you for participating in this survey! We hope that your responses will be very helpful in our research study to make some analysis.

### 2023-03-17 outputs/paper/paper.qmd
Update paper to include minor changes on each figure's axis titles

@@
   geom_bar(aes(x=stress, y = ..prop.., fill=factor(..x..)))+
-  labs(y = "Percent", fill="Work Stress", x = 'Work Stress', title = "Work Stress by status" ) +
+  labs(y = "Percentage", fill="Work Stress", x = 'Work Stress', title = "Work Stress by Status" ) +
   facet_wrap(~wrkstat) +
@@
   geom_bar(aes(x = stress,  y = ..prop.., group = range, fill=factor(..x..))) +
+  labs(y = "Percentage", fill="Work Stress", x = 'Work Stress', title = "Work Stress by Work Hours" ) +
   scale_x_discrete(limits = positions) +
@@
   geom_bar(aes(x = stress,  y = ..prop.., group = wrkgovt, fill=factor(..x..))) +
+  labs(y = "Percentage", fill=" Work Stress", x = 'Work Stress', title = "Work Stress by Industry Sector" ) +
   scale_x_discrete(limits = positions) +
@@
   geom_bar(aes(x = stress,  y = ..prop.., group = satjob, fill=factor(..x..))) +
+  labs(y = "Percentage", fill="Work Stress", x = 'Work Stress', title = "Work Stress by Job Satisfaction" ) +
   scale_x_discrete(limits = positions) +
